### 2015-02-12 paper.Rnw
Added paper first version
First word-vomit of the paper

\documentclass[twoside]{article}
\usepackage{multicol}
\usepackage[hmarginratio=1:1, top=32mm, columnsep=20pt]{geometry}
\usepackage{amsmath}

\author{Sam Helmich}
\title{Board Game Recommendation}

\begin{document}
\maketitle

\section{Introduction}

Board gaming, while being very popular among children, is a niche hobby beyond adolescence. Everyone may have grown up with Monopoly, but most people may have never heard of Citadels or Twilight Imperium. And as such, board gaming is a very difficult hobby to break into. "Gateway Games" such as Settlers of Catan or Ticket to Ride offer a gradual transition between the ultra-popular and mechanics-lite games like monopoly, to more rules-heavy games like Puerto Rico. The most difficult part of the transition, for most people, is actually finding new games to play. New games are often expensive for people who may be apprehensive about trying new games (a copy of Settlers of Catan goes for \$37.80 on Amazon.com). For this reason, a method to recommend games to people would be invaluable.

A few methods already exist, however, their usefulness varies. Boardgamegeek.com, a popular site among frequent board gamers, has a recommendation system that allows users to find a particular game they like, and then from there it will recommend more games that a person may like. A pretty thorough analysis of it and its issues can be found at http://boardgamegeek.com/wiki/page/Game\_Recommendation\_Algorithm. The bigget issue is that it cannot be tailored to an individual. You, an individual user, may have different preferences than the users as a whole, and it might be more useful to tailor recommendations to the individual.

A redditor identified this issue, and created a board game recoomender on the subreddit boardgamerecommender. You can "summon" the recommender by posting your boardgamegeek username anywhere on reddit, preceded by "/u/boardgamerecommender". This will prompt the bot to go search through your ratings of games on boardgamegeek, and then it will make a post with 4 lists of games: New games you may enjoy, older games you may enjoy, games you may find underrated, and games you may find overrated. This is a huge step forward for individuals that are looking for games, as they can now rate games they've played, and can then get recommendations for games they might like. However, there are a few drawbacks. The process has a few possible user inputs (such as playtime, number of players, category, mechanic, etc) that will narrow down the games that will be returned to the user. However, with each specification, a new recommendation will be generated, a process that is said to take 10-30 minutes (FIND A BETTER ESTIMATE FOR THIS). Also, there is no published methodology, so we cannot view how these recommendations are made.

This being said, these are only two board game recommendation systems out of the many out there. In this paper, we seek to build a system the will recommend board games to users based on their own recommendations as well as the recommendations of others.

\section{Data Collection}

data scraped from boardgamegeek.com
scraped in 2 sections - games and ratings
pulled down whole pages due to extensive nature
after pages were pulled down, code was written to extract the pertinent information from each
test data set was dumped into database
database has a few tables - game key, player key, ratings, classifiers, other
the structure of this database allows for easy functionality

\section{Model}
\subsection{Model Selection}
To begin implementing such a model, we can look to two papers on one of the most famous recommender systems of all: Netflix. In 2006, Netflix offered a \$1 Million prize for a improvement on their algorithm for predicting user rating on movies they had not seen. In 2008, two teams: BigChaos and BellKor (the eventual winner), published papers on their methodology. In the end, they ended up taking something of a shotgun approach, using a linear blend of many different models to come up with a single super-model.

There are a few things that make sense to model first. Certain games are more popular than other games, and certain people will naturally be more or less generous than others in their ratings. These two effects are very important to measure, as the first will allow us to provide estimates for people that have rated no games (and therefore we know nothing about), and the second will allow us to fine tune estimates for people who have rated many games. We will implement these in the same way that BigChaos implemented the Global Effects model (page 6, BigChaos paper). While they implemented 14 steps, we will only use the 2, due to the complexity of the calculations required and the time with which it is practical to fit a model.

After we fit the general effects, we wish to model the remaining residuals, ideally in some way that gives us informative numeric summaries about the individual users in the process. In addition, we'd also like to compare a users ratings to other users, and hopefully use the behavior of simiar users to predict new ratings for someone. A tempting first step is to, for a particular game, find other people who have rated that game who have rated games that you've rated. Then, you could find people who rated games similarly to you, and use their ratings of the new game as an estimate. This often doesnt work, as sometimes the number of people who have rated the same games as you can be relatively low, if not 0 (show a graph for this part).

To get around the problem of low crossover in game ratings, we turn to the characteristics of the games themselves. This kills two birds with one stone. We'll create, for each user, a rating for each game mechanic and genre. Then, since there are so many mechanics and genres of game, a player only needs to rate a small, albeit diverse set of games before we can correlate that player's habits with other players, even if they've never played any of the same games! Then we can use those to filter rating results, as we'll know what board game mechanics that a particular player will enjoy.

We now have a very simple two step model that will allow us to predict ratings for new board games that a person may not have even played! Also, it allows us to rate games that nobody has played before: We can use the overall mean as a starting point, then for an individual player, we can adjust the rating based on their ratings of similar games from other genres.

\subsection{Model Fit}
The specifics of the model follow very closely to those outlined in the Global Effects and knn sections of the BigChaos paper [need bib]. Here, we cover the specifics of both.
\subsubsection{Global Effects}
The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference). The general idea is to remove the global mean from all ratings to get a residual for each observation. We'll group these residuals by game, and then find an effect for that game. Subtracting this effect from the residual, we'll take those residuals and repeat the same for each user.

As we proceed $y_{u,i}$ will denote the rating by user $u$ for game $i$. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$

However, we'll approach this in steps, as follows:

\begin{center}
\begin{align}
y_{u,i} &= \mu_{global} + error_{global} \\
error_{global} &= \mu_{i} + error_{game} \\
error_{game} &= \mu_{u} + error_{user} 
\end{align}
\end{center}

And we'll use the estimation method outlined in the paper (need ref) for each, as follows below:

\begin{center}
\begin{align}
\mu_{global} = \frac{\sum_{u = 1}^{N_{u}}\sum_{i = 1}^{N_{games}}}{1}
\end{align}
\end{center}

\section{Implementation}

\subsection{Individualized Summaries}

\subsection{Group Summaries}

\section{Conclusion}

%\begin{multicols}{2}
\section{Data Management}

\section{Model}
Let $y_{ij}$ be the rating for the $i$\textsuperscript{th} player of the $j$\textsuperscript{th} game. We fit the model in two steps, first finding the global residual, and then using a k-means approach to estimation.
\subsection{Global Residual}
The goal of a global residual is to remove two logical biases: some games are naturally rater higher than othes, and certain players are more generous in their ratings than others. The global residual will be the result of removing these baises. Note that this process by itself can be used as a prediction technique, but more importantly, the residuals it generates can be used as the input to other prediction strategies to improve overall performance (as seen in the section where we evaluate the performance of the different strategies). The process for finding these residuals is identical to the method described by [that paper that i used from the netflix part].

\subsection{Model Fitting}
The model fitting needs to be autonomous, so we need to algorithmically proceed through the model fitting. We'll proceed as follows:

\begin{enumerate}
\item Update $\mathbf{\alpha}$
  \begin{enumerate}
  \item Update $\alpha_1$ - parameter for game means
  \item Update $\alpha_2$ - parameter for player means
  \end{enumerate}
\item Update $k$ for k-means
\end{enumerate}

This follows the natural flow of data through the model, so this is the best method for finding parameter estimates.

%\end{multicols}

\end{document}

### 2015-02-19 paper.Rnw
added reference files
quotes and urls

@@
 \usepackage{amsmath}
+\usepackage{natbib}
+\usepackage{url}
+
+\newcommand{\hh}[1]{{\color{magenta} #1}}
 
@@
 And as such, board gaming is a very difficult hobby to break into.
-"Gateway Games" such as Settlers of Catan or Ticket to Ride offer a gradual transition between the ultra-popular and mechanics-lite games like monopoly, to more rules-heavy games like Puerto Rico.
+``Gateway Games" such as Settlers of Catan or Ticket to Ride offer a gradual transition between the ultra-popular and mechanics-lite games like monopoly, to more rules-heavy games like Puerto Rico.
 The most difficult part of the transition, for most people, is actually finding new games to play.
@@
 Boardgamegeek.com, a popular site among frequent board gamers, has a recommendation system that allows users to find a particular game they like, and then from there it will recommend more games that a person may like.
-A pretty thorough analysis of it and its issues can be found at http://boardgamegeek.com/wiki/page/Game\_Recommendation\_Algorithm.
+A pretty thorough analysis of it and its issues can be found at \url{http://boardgamegeek.com/wiki/page/Game_Recommendation_Algorithm}.
 The bigget issue is that it cannot be tailored to an individual.
@@
 A redditor identified this issue, and created a board game recoomender on the subreddit boardgamerecommender.
-You can "summon" the recommender by posting your boardgamegeek username anywhere on reddit, preceded by "/u/boardgamerecommender".
+You can ``summon" the recommender by posting your boardgamegeek username anywhere on reddit, preceded by ``/u/boardgamerecommender".
 This will prompt the bot to go search through your ratings of games on boardgamegeek, and then it will make a post with 4 lists of games: New games you may enjoy, older games you may enjoy, games you may find underrated, and games you may find overrated.
@@
 %\end{multicols}
+\bibliographystyle{asa}
+\bibliography{references}
 

### 2015-02-20 paper.Rnw
Added conflicted changes
Fleshed out some of the modeling section

@@
 The specifics of the model follow very closely to those outlined in the Global Effects and knn sections of the BigChaos paper [need bib]. Here, we cover the specifics of both.
+
+Note for editing: I know this model is a mess at the moment, just getting the general ideas down on paper. 
 \subsubsection{Global Effects}
@@
 
+\subsection{Classifier Aggrigation}
+Independent of the global effects, we need to give each user a score in each classification category.
+We will use 4 types of categorical variables, which will yield (FIND OUT HOW MANY CLASSIFIER VARIABLES).
+To do this, we can either use the ratings themselves, or the residuals from the global effects process.
+Let $c_{u,j}$ be the categorical rating for user $u$ in category $j$.
+Additionally, for each game $i$, there exists a set of classifiers $C_i$ that describe the aspects of the game in things such as mechanics, game family, categories, and subdomains.
+Then for each user, we take all of the categorical values for each user and normalize them, so they should all have mean 0 and standard deviation of 1 for each user.
+Then, within each category, the average rating is calculated.
+Now, if a user hasn't rated anything in this category we can assign them a score of 0 in that category (what should amount to an average rating).
+These serve dual purposes: Firstly, we can assess what sorts of games a user rates higher than average most of the time, and game they rate lower than average most of the time.
+This will allow the algorithm to not only recommend specific games that have a high predicted rating, but will also allow for the user to find categories or mechanics that they enjoy as well.
+The other purpose is that we now have a larger dataset to compare people than we might have had before, as we are guarantedd to have (FIND OUT HOW MANY CLASSIFIER VARIABLES) for every user.
+
+\subsection{k-nearest-neighbors}
+After we aggregate on the classifiers, we can then find correlations between all of the users.
+We will then take the $k$ nearest neighbors, where the distance is given by $\rho_{i,k}$, the correlation between the $i$ and $k$th users.
+In this algorithm, $k$ will be chosen by a crossvalidation procedure that will be done on a user-by-user basis, as determining a global value for $k$ would be far too computationally expensive.
+After the optimal $k$ is found, then for each game the $k$ nearest neighbors' (who have rated the game) rating's are averaged using a weighted mean, with the weights being equal to $\frac{2}{1 + \rho_{i,k}}$.
+
+\subsection{Back to Rating}
+After the k-nearest-neighbors step, we need to "rewind" from the global effects fitting process.
+For each game, we'll need to add on the specific user effect, the effect for each game, and the overall game mean.
+This will then yield a rating for each game for each player!
+
 \section{Implementation}
@@
 
-%\begin{multicols}{2}
-\section{Data Management}
-
-\section{Model}
-Let $y_{ij}$ be the rating for the $i$\textsuperscript{th} player of the $j$\textsuperscript{th} game.
-We fit the model in two steps, first finding the global residual, and then using a k-means approach to estimation.
-\subsection{Global Residual}
-The goal of a global residual is to remove two logical biases: some games are naturally rater higher than othes, and certain players are more generous in their ratings than others.
-The global residual will be the result of removing these baises.
-Note that this process by itself can be used as a prediction technique, but more importantly, the residuals it generates can be used as the input to other prediction strategies to improve overall performance (as seen in the section where we evaluate the performance of the different strategies).
-The process for finding these residuals is identical to the method described by [that paper that i used from the netflix part].
-
-\subsection{Model Fitting}
-The model fitting needs to be autonomous, so we need to algorithmically proceed through the model fitting. We'll proceed as follows:
-
-\begin{enumerate}
-\item Update $\mathbf{\alpha}$
-  \begin{enumerate}
-  \item Update $\alpha_1$ - parameter for game means
-  \item Update $\alpha_2$ - parameter for player means
-  \end{enumerate}
-\item Update $k$ for k-means
-\end{enumerate}
-
-This follows the natural flow of data through the model, so this is the best method for finding parameter estimates.
-
 %\end{multicols}

### 2015-03-02 paper.Rnw
Additions and revisions to paper
Revised the introduction, began the technical implementation. Next up is
work on Bibliography and revising the modeling section, as well as
adding a conclusion.

@@
 ``Gateway Games" such as Settlers of Catan or Ticket to Ride offer a gradual transition between the ultra-popular and mechanics-lite games like monopoly, to more rules-heavy games like Puerto Rico.
-The most difficult part of the transition, for most people, is actually finding new games to play.
-New games are often expensive for people who may be apprehensive about trying new games (a copy of Settlers of Catan goes for \$37.80 on Amazon.com).
+For most people, the difficult part of the transition is actually finding new games to play.
+New games are often expensive for people who may be apprehensive about trying new games (a copy of Settlers of Catan goes for about \$40 on Amazon).
 For this reason, a method to recommend games to people would be invaluable.
 
-A few methods already exist, however, their usefulness varies.
+A few methods of varying usefulness already exist.
 Boardgamegeek.com, a popular site among frequent board gamers, has a recommendation system that allows users to find a particular game they like, and then from there it will recommend more games that a person may like.
-A pretty thorough analysis of it and its issues can be found at \url{http://boardgamegeek.com/wiki/page/Game_Recommendation_Algorithm}.
+A pretty thorough analysis of it (and its issues) can be found at \url{http://boardgamegeek.com/wiki/page/Game_Recommendation_Algorithm}.
 The bigget issue is that it cannot be tailored to an individual.
-You, an individual user, may have different preferences than the users as a whole, and it might be more useful to tailor recommendations to the individual.
+You, an individual user, may have different preferences than the userbase average, and it might be more useful to tailor recommendations to the individual.
 
-A redditor identified this issue, and created a board game recoomender on the subreddit boardgamerecommender.
-You can ``summon" the recommender by posting your boardgamegeek username anywhere on reddit, preceded by ``/u/boardgamerecommender".
-This will prompt the bot to go search through your ratings of games on boardgamegeek, and then it will make a post with 4 lists of games: New games you may enjoy, older games you may enjoy, games you may find underrated, and games you may find overrated.
-This is a huge step forward for individuals that are looking for games, as they can now rate games they've played, and can then get recommendations for games they might like.
-However, there are a few drawbacks.
-The process has a few possible user inputs (such as playtime, number of players, category, mechanic, etc) that will narrow down the games that will be returned to the user.
-However, with each specification, a new recommendation will be generated, a process that is said to take 10-30 minutes (FIND A BETTER ESTIMATE FOR THIS).
-Also, there is no published methodology, so we cannot view how these recommendations are made.
+Recommendations for the individual aren't easy to come by.
+There are a few automated systems out there (such as the boargamerecommender bot on reddit), but for the most part they are time intensive and often lack interactivity or any sort of information that would be useful to the user beyond a few single games that it might recommend.
 
-This being said, these are only two board game recommendation systems out of the many out there.
 In this paper, we seek to build a system the will recommend board games to users based on their own recommendations as well as the recommendations of others.
+This system will offer the user not only recommendations for a very large number of games, but it will also offer insights to the user about what sorts of games they like, not only scores for particular games.
 
@@
 
-data scraped from boardgamegeek.com
-scraped in 2 sections - games and ratings
-pulled down whole pages due to extensive nature
-after pages were pulled down, code was written to extract the pertinent information from each
-test data set was dumped into database
-database has a few tables - game key, player key, ratings, classifiers, other
-the structure of this database allows for easy functionality
+For data, we turn to \url{boardgamegeek.com}, a website that holds information, ratings, and reviews for over 75,000 games.
+Users can create an account and then rate games they've played as well as keep an inventory of games they own.
+All of this information is accessable and scrapable, and will be the data that drives our analysis.
+
+There are two types of particular information that we will be interested in: user ratings of the games and characteristics about the game (referred to as classifiers in the modeling section).
+The characteristics we're interested in fall into these categories:
+
+\begin{center}
+\begin{tabular}{|l|c|}
+\hline
+Category &  This covers the broad categories a game\\
+& might fall into (ex.: Humor, Puzzle, Sports, etc.) \\
+\hline
+Family &  Some games a naturally part of a "family" \\
+&  that share a name or common element \\
+& (ex.: Animals:Bats, Ancient Wars Series, Hello Kitty)\\
+\hline
+Mechanic &  This is what mechanisms the game uses\\
+& (ex.: Dice Rolling, Card Drafting, Worker Placement)\\
+\hline
+Subdomain & More general than Category, but loosely bins games\\
+& (ex.: Family Games, Strategy Games, etc.)\\
+\hline
+\end{tabular}
+\end{center}
+
+The ratings for each game are on a scale from 1 to 10, with a score of 1 indicating a really bad game and a score of 10 being the best possible.
+Each of the ratings are paired with a unique user identification number as well as a game number.
+Combined with the classifiers data, we can provide very specific ratings for each individual user.
 
@@
 
-\subsection{Model Fit}
+\subsection{Model Structure}
 The specifics of the model follow very closely to those outlined in the Global Effects and knn sections of the BigChaos paper [need bib]. Here, we cover the specifics of both.
@@
 
-\subsection{Classifier Aggrigation}
+\subsubsection{Classifier Aggrigation}
 Independent of the global effects, we need to give each user a score in each classification category.
@@
 
-\subsection{k-nearest-neighbors}
+\subsubsection{k-nearest-neighbors}
 After we aggregate on the classifiers, we can then find correlations between all of the users.
@@
 
-\subsection{Back to Rating}
+\subsubsection{Back to Rating}
 After the k-nearest-neighbors step, we need to "rewind" from the global effects fitting process.
@@
 
-\section{Implementation}
+\subsection{Parameter Optimization \& Selection}
+There are three parameters that the model relies on: The two $\alpha$ values that balance out biasedness in the global effects section, and the $k$ in the k-nearest neighbors portion of the algorithm.
+The $\alpha$'s will be done globally (that is, every individual will use the same values for $\alpha_1$ and $\alpha_2$), while the values for $k$ will be done locally (each user will have their own value of $k$).
+While the approaches for finding the optimal values for these parameters may be slightly different, the general implementation will be the same.
+We'll first begin by partitioning a subset of the data into 10 different sections (10 'folds'), and then choosing an appropriate parameter space for the parameter to be optimized.
+Ffor the alpha's, we'll choose values between 0 and 1000 (double check this), evenly spaced, and for $k$ we'll choose a positive integer.
+We'll then proceed fold by fold, using the other 9 folds to fit a model, and then using the data in current fold to check how the model works.
+Using root mean square error, we can then rank each of the possible parameter values in the parameter space, and the value with the lowest rmse will be the 'optimal' value for that parameter.
 
-\subsection{Individualized Summaries}
+\section{Technical Implementation}
+** SCREENSHOTS NEEDED **
 
-\subsection{Group Summaries}
+This algorithm is implemented in R using shiny as a vehhicle for a simple, easy to use user interface.
+Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that will be useful in the classifier aggrigation and k-nearest-neighbors section; the predicted ratings table, which will store predicted ratigs if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user.
+This data storage will allow for ease of implementation: we only really need the ratings table and the classifier table to do all the work, and the key tables are really only needed to display the inforation to the user in a format which they will understand.
+
+In an implementation of this, the application will begin by having the user enter the username they've registered on \url{boardgamegeek.com}.
+Instead of logging the users ratings directly through the application, we will instead leverage the existing infrastructure of boardgame geek.
+After checking that its a valid username, the application will scour boardgamegeek to see if the database's information is complete: that is, do we have this user's data in the player key?
+Have they rated games that aren't in our database?
+Have they rerated games that are in our database?
+Have they rated games that aren't in our database?
+.If so, we also need to bring in other users ratings for any games that we may not have had in our database (since a single rating for a game isn't very useful and will greatly skew the results for that game).
+After this initial step and database update, we can see if they already have saved predicted ratings in the system, and if not, we'll proceed to model fitting.
+
+We first apply the global effects strategy as descibed in the modling section.
+The optimization of the $\alpha$ parameters will have occred sometime before this and the algoritm will use use those stored values.
+A weeky or monthly automatic updaing procedure is used to keep these values up to date as more data comes into the system.
+From the global effects, we then aggregate those based on the classifier aggrigation procedure as described above.
+These are then stored in the database for future use by the system.
+After that, before we move onto the k-nearest neighbors implementation, we must first find an optimal value for $k$ using the 10-fold cross validation as described above.
+
+Finally, after the k-fold crossvalidation, we're ready to run the other games through the algorithm to get a predicted score.
+The games that the user has already rated will then be replaced with those ratings, and the full list will be outputted for the user.
+
+Since we saved the aggregated classifier ratings for the individual user, we can suggest ways for the user to subset this data.
+By providing the graph below to the player (GET THE GRAPH WORKING), they can see for themselves their not-necessarily obvious preferences in game family or mechanic.
+They can then search by their preferred mechanic or theme, and get ratings for games that fit that description.
+Or the user can simply sort by highest rated.
 
@@
 
-%\end{multicols}
 \bibliographystyle{asa}
 \bibliography{references}
+\def\thebibliography#1{\section*{References\@mkboth
+ {REFERENCES}{REFERENCES}}\list
+%% default is no labels, for those not using \cite or BibTeX
+% undent first line of a reference by using negative \itemindent.
+ {}{\setlength{\leftmargin}{3em}\setlength{\labelsep}{0pt}\itemindent=-\leftmargin}
+ \def\newblock{\hskip .11em plus .33em minus -.07em}
+ \sloppy\clubpenalty4000\widowpenalty4000
+ \sfcode`\.=1000\relax}
+
+\def\@biblabel#1{\hfill}
+
+% don't box citations, add space between multiple citations, separate with ;
+\def\@citex[#1]#2{\if@filesw\immediate\write\@auxout{\string\citation{#2}}\fi
+  \def\@citea{}\@cite{\@for\@citeb:=#2\do
+    {\@citea\def\@citea{; }\@ifundefined
+       {b@\@citeb}{{\bf ?}\@warning
+       {Citation `\@citeb' on page \thepage \space undefined}}%
+{\csname b@\@citeb\endcsname}}}{#1}}
 

### 2015-03-02 paper.Rnw
couple of suggestions

@@
 There are two types of particular information that we will be interested in: user ratings of the games and characteristics about the game (referred to as classifiers in the modeling section).
-The characteristics we're interested in fall into these categories:
+The characteristics we're interested in fall into these categories as outlined in table~\ref{tab:categories}
 
-\begin{center}
+\hh{Just to make this flow better - make the table below an actual table in the table environment. Give it a caption and refer to it in the text ... I've just implemented it for you}
+\begin{table}[ht]
+\centering
 \begin{tabular}{|l|c|}
@@
 \end{tabular}
-\end{center}
+\caption{\label{tab:categories} Overview of the  relevant characteristics of board games considered in this project.}
+\end{table}
 
@@
 Combined with the classifiers data, we can provide very specific ratings for each individual user.
+
+\hh{Give an example of how the data that you have scraped looks like. This would be a good point to introduce a bit of the notation that you are using below, like the $y_{u,i}$. }
+
+\hh{Aren't you continuously scraping? You should mention this at this point, and also talk a bit about how you have implemented the scraping.}
 
@@
 These two effects are very important to measure, as the first will allow us to provide estimates for people that have rated no games (and therefore we know nothing about), and the second will allow us to fine tune estimates for people who have rated many games.
-We will implement these in the same way that BigChaos implemented the Global Effects model (page 6, BigChaos paper).
-While they implemented 14 steps, we will only use the 2, due to the complexity of the calculations required and the time with which it is practical to fit a model.
+We will implement these in the same way that BigChaos implemented the Global Effects model (page 6, BigChaos paper \hh{which of the BigChaos papers are you refering to here, exactly?
+At some point you need to include the exact reference in the bib tex, but I'd just like to know for myself ...}).
+While they implemented 14 steps, we will only use the 2 \hh{which 2?
+you need to include a few more details here}, due to the complexity of the calculations required and the time with which it is practical to fit a model.
 
@@
 Note for editing: I know this model is a mess at the moment, just getting the general ideas down on paper. 
+\hh{this is not nearly as messy as you make it out to be, but you need to make sure to define every little piece of notation as you, because otherwise your reader will be lost.}
 \subsubsection{Global Effects}

### 2015-03-02 paper.Rnw
couple more comments

@@
 \subsubsection{Global Effects}
-The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference).
  The general idea is to remove the global mean from all ratings to get a residual for each observation.
@@
 Subtracting this effect from the residual, we'll take those residuals and repeat the same for each user.
+The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference, \hh{you still need to paraphrase the motivation here.}).
 
-As we proceed $y_{u,i}$ will denote the rating by user $u$ for game $i$. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$
+As we proceed $y_{u,i}$ will denote the rating by user $u$ for game $i$ \hh{what scale is the rating on?}. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$
 
-However, we'll approach this in steps, as follows:
+However, we'll approach this in steps, as follows: \hh{what is the benefit of using a hierarchical approach?}
 
@@
 \end{center}
+\hh{something went wrong in the formula - what are you summing over?}
 
-\subsubsection{Classifier Aggrigation}
+\hh{This is a linear approach to estimate ratings -- one of the problems is, that rating is on a finite scale, i.e.
+it is possible that predictions go outside this interval - what do you do with those predictions?}
+
+\subsubsection{Classifier Aggregation}
 Independent of the global effects, we need to give each user a score in each classification category.
@@
 This algorithm is implemented in R using shiny as a vehhicle for a simple, easy to use user interface.
-Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that will be useful in the classifier aggrigation and k-nearest-neighbors section; the predicted ratings table, which will store predicted ratigs if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user.
+Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that will be useful in the classifier aggregation and k-nearest-neighbors section; the predicted ratings table, which will store predicted ratigs if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user.
 This data storage will allow for ease of implementation: we only really need the ratings table and the classifier table to do all the work, and the key tables are really only needed to display the inforation to the user in a format which they will understand.
@@
 A weeky or monthly automatic updaing procedure is used to keep these values up to date as more data comes into the system.
-From the global effects, we then aggregate those based on the classifier aggrigation procedure as described above.
+From the global effects, we then aggregate those based on the classifier aggregation procedure as described above.
 These are then stored in the database for future use by the system.

### 2015-03-05 paper.Rnw
Revisions and Additions
Revised the modeling section, added some data for some tables

@@
 \newcommand{\hh}[1]{{\color{magenta} #1}}
+\newcommand{\sam}[1]{{\color{blue} #1}}
 
@@
 
+<<loader, echo = FALSE>>=
+library(xtable)
+load("paper_stuff")
+@
 \section{Introduction}
@@
 There are two types of particular information that we will be interested in: user ratings of the games and characteristics about the game (referred to as classifiers in the modeling section).
-The characteristics we're interested in fall into these categories as outlined in table~\ref{tab:categories}
+The characteristics we're interested in fall into these categories as outlined in table~%\ref{tab:categories}
 
@@
 
+<<classifiers, echo = FALSE, results = 'asis'>>=
+xtable(paper_stuff$classifiers, caption = c("Classifiers Data, as seen in Database", "Classifiers Table"))
+@
+
 The ratings for each game are on a scale from 1 to 10, with a score of 1 indicating a really bad game and a score of 10 being the best possible.
@@
 Combined with the classifiers data, we can provide very specific ratings for each individual user.
+<<ratings, echo = FALSE, results = 'asis'>>=
+xtable(paper_stuff$ratings, caption = c("Ratings Data, as seen in Database", "Ratings Table"))
+@
+
+To keep the information as compact as possible, individual players and games are referred to by an ID number, which corresponds to a row in one of the key tables.
+These key tables contain information about the players and game that we don't necessarily need to include with each rating or classifier.
+This greatly reduces the size of the database needed to store and work with the information.
+<<player_key, echo = FALSE, results = 'asis'>>=
+xtable(paper_stuff$player_key, caption = c("The Player Key relates important info about the individual to the userid", "Player Key Table"))
+@
+
+<<game_key, echo = FALSE, results = 'asis'>>=
+xtable(paper_stuff$game_key[,1:6], caption = c("The Game Key gives useful information about the game", "Game Key Table"))
+@
+
+Not all of the data was scraped before creating the applicaiton.
+Data for 6,088 games was scraped and stored across the four tables.
+The reason for not scraping all of the data is twofold.
+First of all, for those 6,088 games there are over 3 million unique ratings.
+If we were to expand it to all 75,756 games known to BoardGameGeek, then our data set would be crippilingly large for the small desktop I have to work on.
+Secondly, this will force us to implement a continuous data scraping procedure.
+Each time the person uses the app, all we have to do is make sure we have all of their current information correct.
+We look through our ratings and can check them with their userpage on BoardGameGeek.
+From this, we can revise old ratings or add new ones, as well as check to see that all of the games that they've rated are in our database!
 
 \hh{Give an example of how the data that you have scraped looks like. This would be a good point to introduce a bit of the notation that you are using below, like the $y_{u,i}$. }
-
-\hh{Aren't you continuously scraping? You should mention this at this point, and also talk a bit about how you have implemented the scraping.}
+\sam{I still need to tie this back to the modeling section.}
 
@@
 \subsubsection{Global Effects}
- The general idea is to remove the global mean from all ratings to get a residual for each observation.
+The general idea behind global effects is to account for the natural biases present in particular games or players.
+For example, we might have two players who both play and enjoy the same game the same amount.
+However, one person might give this game a score of 10 while the other gives it a score of 9.
+Despite their overall enjoyment being the same, perhaps the latter player just tends to be a harsher rater in general.
+The global effect for this person would then be lower than that of the much more cheery fellow who gave the game a 10.
+Similarly, this effect can be present for games due to any number of reasons.
+
+We'll estimate these effects one at a time using a heirarchical structure. After estimating the effect, we'll be left with a residual, which will be used as the response for the next level.
+The global effects formual is
+
+\begin{equation}
+r_{u,i} = \theta_u x_{u,i} + error
+\end{equation}
+
+Where for the first step, $r_{u,i}$ will be the rating from user $u$ for game $i$ and $x_{u,i}$ will be an indicator variable that is 1 if player $u$ has rated game $i$ and 0 otherwise.
+In the second step $r_{u,i}$ will be the residual from the first portion, the subscript on $\theta$ becomes $i$ and $x_{u,i}$ will be an indicator variable that is 1 if the rating is for game $i$ and 0 otherwise.
+
+To get an estimate of $\theta$ at either step, we'll use the average of the nonzero values which are shunk by the fitting parameter $\alpha$ to avoid overfitting.
+The sum is over either all ratings for user $u$ or all ratings for game $i$, depending on whether you're estimating $\theta_u$ or $\theta_i$.
+Either way, the structure is the same.
+\begin{equation}
+\hat{\theta_u} = \frac{\sum r_{u,i}}{n_u + \alpha}
+\end{equation}
+
+%  The general idea is to remove the global mean from all ratings to get a residual for each observation.
 We'll group these residuals by game, and then find an effect for that game.
@@
 The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference, \hh{you still need to paraphrase the motivation here.}).
-
-As we proceed $y_{u,i}$ will denote the rating by user $u$ for game $i$ \hh{what scale is the rating on?}. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$
-
-However, we'll approach this in steps, as follows: \hh{what is the benefit of using a hierarchical approach?}
-
-\begin{center}
-\begin{align}
-y_{u,i} &= \mu_{global} + error_{global} \\
-error_{global} &= \mu_{i} + error_{game} \\
-error_{game} &= \mu_{u} + error_{user} 
-\end{align}
-\end{center}
-
-And we'll use the estimation method outlined in the paper (need ref) for each, as follows below:
-
-\begin{center}
-\begin{align}
-\mu_{global} = \frac{\sum_{u = 1}^{N_{u}}\sum_{i = 1}^{N_{games}}}{1}
-\end{align}
-\end{center}
-\hh{something went wrong in the formula - what are you summing over?}
+% 
+% To begin $r_{u,i}$ will denote the rating by user $u$ for game $i$, where $y_{u,i}$ is a number between 1 and 10. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$
+% 
+% However, we'll approach this in steps, as follows: \hh{what is the benefit of using a hierarchical approach?}
+% 
+% \begin{center}
+% \begin{align}
+% y_{u,i} &= \mu_{global} + error_{global} \\
+% error_{global} &= \mu_{i} + error_{game} \\
+% error_{game} &= \mu_{u} + error_{user} 
+% \end{align}
+% \end{center}
+% 
+% And we'll use the estimation method outlined in the paper (need ref) for each, as follows below:
+% 
+% \begin{center}
+% \begin{align}
+% \mu_{global} = \frac{\sum_{u = 1}^{N_{u}}\sum_{i = 1}^{N_{games}}}{1}
+% \end{align}
+% \end{center}
+% \hh{something went wrong in the formula - what are you summing over?}
 

### 2015-03-10 paper.Rnw
Added some bibliography and some clarifications
Added some bib and some clarification on some sections. Most work has
been touching up the application.

@@
 
-\hh{Give an example of how the data that you have scraped looks like. This would be a good point to introduce a bit of the notation that you are using below, like the $y_{u,i}$. }
-\sam{I still need to tie this back to the modeling section.}
-
 \section{Model}
@@
 These two effects are very important to measure, as the first will allow us to provide estimates for people that have rated no games (and therefore we know nothing about), and the second will allow us to fine tune estimates for people who have rated many games.
-We will implement these in the same way that BigChaos implemented the Global Effects model (page 6, BigChaos paper \hh{which of the BigChaos papers are you refering to here, exactly?
-At some point you need to include the exact reference in the bib tex, but I'd just like to know for myself ...}).
-While they implemented 14 steps, we will only use the 2 \hh{which 2?
-you need to include a few more details here}, due to the complexity of the calculations required and the time with which it is practical to fit a model.
+We will implement these in a similar manner to the way that BigChaos implemented their Global Effects model \citep{bigchaos}.
+While they implemented 14 steps including Movie Effect, User Effect, and those variables crossed with variables like time since rating, average rating for the movie, production year, standard deviations, as well as others.
+From the RMSE of their probe set, the difference between using only user/movie effcts and using all 14 effects was only .02, a decrease of about 2\%.
+Noting this negligible increase in accuracy and recognizing my lack of significant computing resources, we will only use user and game effects.
 
@@
 
-To get an estimate of $\theta$ at either step, we'll use the average of the nonzero values which are shunk by the fitting parameter $\alpha$ to avoid overfitting.
+To get an estimate of $\theta$ at either step, we'll use the average of the nonzero values which are shunk by the fitting parameter $\alpha$ to avoid overfitting (motivation for using $\alpha$ in this manner can be found in \cite{scalecollab}).
 The sum is over either all ratings for user $u$ or all ratings for game $i$, depending on whether you're estimating $\theta_u$ or $\theta_i$.
@@
 
+Noting that this can produce esimates that are above 10 and below 1, ratings that fall outside of $[1,10]$ are truncated to the closes interval endpoint.
 %  The general idea is to remove the global mean from all ratings to get a residual for each observation.
@@
 % \hh{something went wrong in the formula - what are you summing over?}
-
-\hh{This is a linear approach to estimate ratings -- one of the problems is, that rating is on a finite scale, i.e.
-it is possible that predictions go outside this interval - what do you do with those predictions?}
 

### 2015-03-19 paper.Rnw
Added parameter plots
Added parameter plots and added to the parameter optimization section.
Screenshots forthcoming

@@
 The $\alpha$'s will be done globally (that is, every individual will use the same values for $\alpha_1$ and $\alpha_2$), while the values for $k$ will be done locally (each user will have their own value of $k$).
+
+\subsubsection{Optimizing $\alpha$}
+To optimize the $\alpha$ values, we'll take all of the ratings for all players and subtract off the mean of all of the ratings (this is the first step in the globalized residual process).
+This leaves us with three variables: The userid, the gameid and the residual.
+We'll randomly split all of the data into 10 equally sized buckets.
+We'll use each of these buckets as a test group, where we put the other buckets together, fit a model using a canidate value for $\alpha$, and then fit the values in the test group.
+From each test group, we'll then find the mean square error, and then average those together from all 10 groups.
+Taking the square root of that average, we can get the RMSE for each canidate value for $\alpha$.
+Finally, we'll proceed using the canidate value for $\alpha$ with the lowest RMSE.
+A graph of this process can be seen below, with the domain trunacted for easier visibility.
+
+<<game_alpha_optim, echo = F>>=
+load("game_graph")
+game_graph
+@
+
+<<player_alpha_optim, echo = F>>=
+load("player_graph")
+player_graph
+@
+
+<<k_optim, echo = F>>=
+load("k_plot")
+k_plot
+@
+
 While the approaches for finding the optimal values for these parameters may be slightly different, the general implementation will be the same.

### 2015-03-19 paper.Rnw
couple of suggestions, read-through section 3.2

@@
 \usepackage{url}
+\usepackage[normalem]{ulem}
 
@@
 \maketitle
+\begin{abstract}
+\hh{You'll need to include an abstract here}
+\end{abstract}
 
@@
 There are two types of particular information that we will be interested in: user ratings of the games and characteristics about the game (referred to as classifiers in the modeling section).
-The characteristics we're interested in fall into these categories as outlined in table~%\ref{tab:categories}
+The characteristics we're interested in fall into these categories as outlined in table~\ref{tab:categories}.
 
-\hh{Just to make this flow better - make the table below an actual table in the table environment. Give it a caption and refer to it in the text ... I've just implemented it for you}
 \begin{table}[ht]
@@
 
+\hh{Refer to table~\ref{tab:classifiers} somewhere in the text, add a label to the ratings table as well.}
 <<classifiers, echo = FALSE, results = 'asis'>>=
-xtable(paper_stuff$classifiers, caption = c("Classifiers Data, as seen in Database", "Classifiers Table"))
+print(xtable(paper_stuff$classifiers, caption = c("Example lines of classifiers data, as seen in Database", "Classifiers Table"), label="tab:classifiers"), include.rownames =FALSE )
 @
@@
 First of all, for those 6,088 games there are over 3 million unique ratings.
-If we were to expand it to all 75,756 games known to BoardGameGeek, then our data set would be crippilingly large for the small desktop I have to work on.
-Secondly, this will force us to implement a continuous data scraping procedure.
-Each time the person uses the app, all we have to do is make sure we have all of their current information correct.
-We look through our ratings and can check them with their userpage on BoardGameGeek.
-From this, we can revise old ratings or add new ones, as well as check to see that all of the games that they've rated are in our database!
+If we were to expand it to all 75,756 games known to BoardGameGeek, then our data set would be crippingly large for all but very high-powered machines.
+%the small desktop I have to work on.
+\hh{Instead, we} %Secondly, this will force us to 
+implement a continuous data scraping procedure \hh{and update the recommender model dynamically:} Each time a person uses the app, %all we have to do is 
+we make sure to check that all of their current information is correct.
+We look through known ratings and  scrape  their userpage on BoardGameGeek.
+Based on this, we  revise old ratings and add new ones, as well as check to see that all of the games that this user has rated are included in the database!
 
@@
 \subsection{Model Selection}
-To begin implementing such a model, we can look to two papers on one of the most famous recommender systems of all: Netflix.
-In 2006, Netflix offered a \$1 Million prize for a improvement on their algorithm for predicting user rating on movies they had not seen.
+\hh{We base our implementation of a recommender system} %To begin implementing such a model, we can look to 
+on two papers  of on one the most famous recommender systems of all: Netflix.
+In 2006, Netflix offered a \$1 Million prize for an improvement on their algorithm for predicting user rating on movies they had not seen.
 In 2008, two teams: BigChaos and BellKor (the eventual winner), published papers on their methodology.
-In the end, they ended up taking something of a shotgun approach, using a linear blend of many different models to come up with a single super-model.
+In the end, these teams ended up taking something of a shotgun approach, using a linear blend of many different models to come up with a single super-model.
+
+\hh{XXX What is the approach you are taking compared to Netflix? the same? similar? what is different? - One or two sentences here.}
 
@@
 These two effects are very important to measure, as the first will allow us to provide estimates for people that have rated no games (and therefore we know nothing about), and the second will allow us to fine tune estimates for people who have rated many games.
-We will implement these in a similar manner to the way that BigChaos implemented their Global Effects model \citep{bigchaos}.
+We will implement these in a similar manner to the way that BigChaos implemented the Global Effects model \citep{bigchaos}.
+\hh{XXX the next clause needs an end to be a sentence :)}
 While they implemented 14 steps including Movie Effect, User Effect, and those variables crossed with variables like time since rating, average rating for the movie, production year, standard deviations, as well as others.
+\hh{XXX explain RMSE; make sure in the next paragraph to always point out what is yours and what is netflix. You know all those distinctions, but for any other reader it is not that obvious.}
 From the RMSE of their probe set, the difference between using only user/movie effcts and using all 14 effects was only .02, a decrease of about 2\%.
@@
 
-Where for the first step, $r_{u,i}$ will be the rating from user $u$ for game $i$ and $x_{u,i}$ will be an indicator variable that is 1 if player $u$ has rated game $i$ and 0 otherwise.
+\hh{XXX what do you mean by first step? If you need iterations, use $r_{u,i}^{(n)}$ or something similar. If there is an iteration, make sure to give a convergence criterion.
+
+Ah, ok - I see - I think what is going on is that you first get rid of the individuals' effects, and then of the games' effect.
+Write that out in two separate pieces.
+Give them appropriate names :)  - then you will have a much easier time describing what you are doing than in the paragraph below, where you are switching out us and is.
+}
+\hh{XXX: doesn't this error need to be indexed as well? could we call the error $\epsilon$ and make away with the text in the formula? }
+\hh{XXX: generally:  no future tense in the write up. The CC is in the here and now :)
+}
+
+Where for the first step, $r_{u,i}$ \sout{will be} \hh{is} the rating from user $u$ for game $i$ and $x_{u,i}$ will be an indicator variable that is 1 if player $u$ has rated game $i$ and 0 otherwise.
 In the second step $r_{u,i}$ will be the residual from the first portion, the subscript on $\theta$ becomes $i$ and $x_{u,i}$ will be an indicator variable that is 1 if the rating is for game $i$ and 0 otherwise.
 
-To get an estimate of $\theta$ at either step, we'll use the average of the nonzero values which are shunk by the fitting parameter $\alpha$ to avoid overfitting (motivation for using $\alpha$ in this manner can be found in \cite{scalecollab}).
+\hh{XXX it is $\theta_u$, not $\theta$. }
+\hh{XXX Why do you call $\alpha$ a fitting parameter and not a shrinking parameter?}
+
+To get an estimate of $\theta$ at either step, we'll use the average of the nonzero values which are shrunk by the fitting parameter $\alpha$ to avoid overfitting (motivation for using $\alpha$ in this manner can be found in \cite{scalecollab}).
 The sum is over either all ratings for user $u$ or all ratings for game $i$, depending on whether you're estimating $\theta_u$ or $\theta_i$.
 Either way, the structure is the same.
+\hh{Again, write this out in two separate formulae.}
 \begin{equation}
-\hat{\theta_u} = \frac{\sum r_{u,i}}{n_u + \alpha}
+\widehat{\theta_u} = \frac{\sum r_{u,i}}{n_u + \alpha}
 \end{equation}
 
-Noting that this can produce esimates that are above 10 and below 1, ratings that fall outside of $[1,10]$ are truncated to the closes interval endpoint.
+Noting that this can produce esimates that are above 10 and below 1, ratings that fall outside of $[1,10]$ are truncated to the closest interval endpoint.
 %  The general idea is to remove the global mean from all ratings to get a residual for each observation.
@@
 Finally, we'll proceed using the canidate value for $\alpha$ with the lowest RMSE.
-A graph of this process can be seen below, with the domain trunacted for easier visibility.
+A graph of this process can be seen \hh{in figure~\ref{fig:alpha}}, with the domain truncated for easier visibility.
 
+\begin{figure}
 <<game_alpha_optim, echo = F>>=
@@
 @
+\caption{\label{fig:alpha} You need a caption here - what is shown?}
+\end{figure}
 

### 2015-03-26 paper.Rnw
Revised the modeling section
Revised the modeling section up to and including the parameter
optimization section. Technical implementation awaiting screenshots. I'm
digging for a database error somewhere in the middle of the app.

@@
 
-We'll estimate these effects one at a time using a heirarchical structure. After estimating the effect, we'll be left with a residual, which will be used as the response for the next level.
-The global effects formual is
+We'll estimate these effects one at a time using a heirarchical structure.
+Define $r_{u,i}^{(1)}$ be the rating from user $u$ for game $i$.
+The first effect we'll fit is a global mean, which we'll denote with the parameter $\theta_g$, and we'll model the ratings using the following formula:
 
 \begin{equation}
-r_{u,i} = \theta_u x_{u,i} + error
+r_{u,i}^{(1)} = \theta_g + \epsilon_{u,i}^{(1)}
 \end{equation}
 
-\hh{XXX what do you mean by first step? If you need iterations, use $r_{u,i}^{(n)}$ or something similar. If there is an iteration, make sure to give a convergence criterion.
+$\theta_g$ is estimated using a simple mean of the $r_{u,i}^{(1)}$'s.
+After removing a global mean, we'll model the global effect for each game using the residuals from the global mean step.
+For this purpose, define $r_{u,i}^{(2)} = \epsilon_{u,i}^{(1)}$, which we'll model using global effect for each game, denoted $\theta_i$.
+The residuals are then modeled using the following, where $x_i$ is an indicator varialbe that the residual belongs to game $i$:
 
-Ah, ok - I see - I think what is going on is that you first get rid of the individuals' effects, and then of the games' effect.
-Write that out in two separate pieces.
-Give them appropriate names :)  - then you will have a much easier time describing what you are doing than in the paragraph below, where you are switching out us and is.
-}
-\hh{XXX: doesn't this error need to be indexed as well? could we call the error $\epsilon$ and make away with the text in the formula? }
+\begin{equation}
+r_{u,i}^{(2)} = \theta_i x_{i} + \epsilon_{u,i}^{(2)}
+\end{equation}
+
+To estimate $\theta_i$, we want something that is like a mean, but also contains an offset to avoid overfitting.
+Following from \cite{scalecollab}, $\widehat{\theta}_{i}$ is estimated using the following, where $\alpha_{game}$ is a tuning parameter, and the sum is over all residuals for game $i$ and $n_{i}$ is the number of residuals for game $i$:
+
+\begin{equation}
+\widehat{\theta}_{i} = \frac{\sum r_{u,i}^{(2)}}{n_{i} + \alpha_{game}}
+\end{equation}
+
+The final step in the global effects stage is finding the global effect for each user denoted $\theta_{u}$.
+In a simillar manner to the game global effects, once again taking the residual from the previous step and model them.
+To proceed, let $r_{u,i}^{(3)} = \epsilon_{u,i}^{(2)}$, and those residuals are modeled using:
+
+\begin{equation}
+r_{u,i}^{(3)} = \theta_{u}x_u + \epsilon_{u,i}^{(3)}
+\end{equation}
+
+And again, $\theta_{u}$ is estimated with $\widehat{\theta}_{u}$, which is calculated with:
+
+\begin{equation}
+\widehat{\theta}_{u} = \frac{\sum r_{u,i}^{(2)}}{n_{i} + \alpha_{player}}
+\end{equation}
+
+Where $\alpha_{player}$ is a tuning parameter to prevent overfitting.
+
 \hh{XXX: generally:  no future tense in the write up. The CC is in the here and now :)
@@
 
-Where for the first step, $r_{u,i}$ \sout{will be} \hh{is} the rating from user $u$ for game $i$ and $x_{u,i}$ will be an indicator variable that is 1 if player $u$ has rated game $i$ and 0 otherwise.
-In the second step $r_{u,i}$ will be the residual from the first portion, the subscript on $\theta$ becomes $i$ and $x_{u,i}$ will be an indicator variable that is 1 if the rating is for game $i$ and 0 otherwise.
-
 \hh{XXX it is $\theta_u$, not $\theta$. }
 \hh{XXX Why do you call $\alpha$ a fitting parameter and not a shrinking parameter?}
-
-To get an estimate of $\theta$ at either step, we'll use the average of the nonzero values which are shrunk by the fitting parameter $\alpha$ to avoid overfitting (motivation for using $\alpha$ in this manner can be found in \cite{scalecollab}).
-The sum is over either all ratings for user $u$ or all ratings for game $i$, depending on whether you're estimating $\theta_u$ or $\theta_i$.
-Either way, the structure is the same.
-\hh{Again, write this out in two separate formulae.}
-\begin{equation}
-\widehat{\theta_u} = \frac{\sum r_{u,i}}{n_u + \alpha}
-\end{equation}
 
@@
 \subsubsection{Classifier Aggregation}
-Independent of the global effects, we need to give each user a score in each classification category.
-We will use 4 types of categorical variables, which will yield (FIND OUT HOW MANY CLASSIFIER VARIABLES).
-To do this, we can either use the ratings themselves, or the residuals from the global effects process.
-Let $c_{u,j}$ be the categorical rating for user $u$ in category $j$.
-Additionally, for each game $i$, there exists a set of classifiers $C_i$ that describe the aspects of the game in things such as mechanics, game family, categories, and subdomains.
-Then for each user, we take all of the categorical values for each user and normalize them, so they should all have mean 0 and standard deviation of 1 for each user.
-Then, within each category, the average rating is calculated.
-Now, if a user hasn't rated anything in this category we can assign them a score of 0 in that category (what should amount to an average rating).
-These serve dual purposes: Firstly, we can assess what sorts of games a user rates higher than average most of the time, and game they rate lower than average most of the time.
-This will allow the algorithm to not only recommend specific games that have a high predicted rating, but will also allow for the user to find categories or mechanics that they enjoy as well.
-The other purpose is that we now have a larger dataset to compare people than we might have had before, as we are guarantedd to have (FIND OUT HOW MANY CLASSIFIER VARIABLES) for every user.
+After finding the global effects for the games and the users, we're left with residuals for every rating that every user has made.
+These residuals represent the last of what differentiate the users based on their preferences for the nuances of each game.
+We need a way to quantify these nuances, and then attempt to model these residual so we can get a better understanding of each individual user.
+Each game in the dataset can be described by things like what mechanics it has, or what family of game it falls under, when it was published, etc.
+All in all, there are up to 1,851 different quantifiable ways that were chosed to descibe games.
+For example, the game Dragonmaster was published in 1981, has a playingtime of 30 minutes, is under the family "Animals: Dragons", falls under the categories "Card Game", "Fantasy Game", is in the subdomain "Strategy Games", uses the mechanic "Trick-taking", and seats 3 or 4 players.
+Each one of these will be treated as a classifier for the game.
+
+In order to model the remaining residuals, we'll give each player a score based on how they rated games with similar classifiers; that is, how did they rate games with the "Trick-taking" mechanic, or how did they rate "Card Games"?
+
+Define $C_{u,z}$ to the be the classifier score for user $u$ for classifier $z$ where $z$ is an index of the 1,851 classifiers. Then:
+
+\begin{equation}
+C_{u,z} = \frac{\sum \epsilon_{u,i \in z}^{(3)}}{n_{z}}
+\end{equation}
+
+where the sum is over all residuals that belong to a game that can be described by classifier $z$.
+For example, if the classifier was the mechanic "Trick-taking" and the user was $u = 3$, the classifier score would be the mean of all of the $\epsilon_{3,i}^{(3)}$ for \textit{all} games that the user has rated that have the "Trick-taking" mechanic.
+If its the case that a user has rated no games with a particular classifier, we'll leave it as an unobserved value.
 
 \subsubsection{k-nearest-neighbors}
-After we aggregate on the classifiers, we can then find correlations between all of the users.
-We will then take the $k$ nearest neighbors, where the distance is given by $\rho_{i,k}$, the correlation between the $i$ and $k$th users.
-In this algorithm, $k$ will be chosen by a crossvalidation procedure that will be done on a user-by-user basis, as determining a global value for $k$ would be far too computationally expensive.
-After the optimal $k$ is found, then for each game the $k$ nearest neighbors' (who have rated the game) rating's are averaged using a weighted mean, with the weights being equal to $\frac{2}{1 + \rho_{i,k}}$.
+Now we have a few classifier values defined for each user who has rated at least one game.
+In order to create predictions for other games that the user hasn't played, we'll find the correlation between players based on their classifier values.
+For each pair of players, the correlation between them can be found by finding all of the classifiers that have scores for both players, and then finding the correlation between those values.
+To make sure that we don't find high correlations between players who only have very few classifier scores in common, we'll on consider a correlation between players if they have more than 6 common classifier scores.
+
+Now to generate an estimate for user $u$ for game $i$, we'll find all of the players who have a correlation value with user $u$ who have also rated game $i$.
+The rating for game $i$ is then the weighted average of the $k$ most highly correlated (or nearest) neighbors of user $u$.
+
+The weights for this will be $\frac{1}{1 + \rho_{u,k}}$ where $\rho_{u,k}$ is the correlation between user $u$ and user $k$.
+The number of neighbors to use $k$ will be determined by cross-validation on a user-to-user basis.
+In the end $z_{u,i}$ will be the k-nearest-neighbors estimate for user $u$ and game $i$, as given in the equation below (where $u_{[j]}$ is the user with the $j^{\text{th}}$ highest correlation between themselves and user $u$:
+
+\begin{equation}
+z_{u,i} = \frac{\sum_{j = 1}^{k} \epsilon_{u_{[j]},i}^{(3)} * \frac{1}{1 + \rho_{u,u_{[j]}}}}{\sum_{j = 1}^{k}\frac{1}{1 + \rho_{u,u_{[j]}}}}
+\end{equation}
 
 \subsubsection{Back to Rating}
-After the k-nearest-neighbors step, we need to "rewind" from the global effects fitting process.
-For each game, we'll need to add on the specific user effect, the effect for each game, and the overall game mean.
-This will then yield a rating for each game for each player!
+From what we've produced above, of the estimated rating for user $u$ and game $i$ is given by:
+
+\begin{equation}
+\widehat{r}_{u,i} = \widehat{\theta}_g + \widehat{\theta}_{i} + \widehat{\theta}_{u} + z_{u,i}
+\end{equation}
 
@@
 There are three parameters that the model relies on: The two $\alpha$ values that balance out biasedness in the global effects section, and the $k$ in the k-nearest neighbors portion of the algorithm.
-The $\alpha$'s will be done globally (that is, every individual will use the same values for $\alpha_1$ and $\alpha_2$), while the values for $k$ will be done locally (each user will have their own value of $k$).
+The $\alpha$'s will be done globally (that is, every individual will use the same values for $\alpha_{game}$ and $\alpha_{player}$), while the values for $k$ will be done locally (each user will have their own value of $k$).
 
 \subsubsection{Optimizing $\alpha$}
-To optimize the $\alpha$ values, we'll take all of the ratings for all players and subtract off the mean of all of the ratings (this is the first step in the globalized residual process).
-This leaves us with three variables: The userid, the gameid and the residual.
-We'll randomly split all of the data into 10 equally sized buckets.
-We'll use each of these buckets as a test group, where we put the other buckets together, fit a model using a canidate value for $\alpha$, and then fit the values in the test group.
-From each test group, we'll then find the mean square error, and then average those together from all 10 groups.
-Taking the square root of that average, we can get the RMSE for each canidate value for $\alpha$.
-Finally, we'll proceed using the canidate value for $\alpha$ with the lowest RMSE.
-A graph of this process can be seen \hh{in figure~\ref{fig:alpha}}, with the domain truncated for easier visibility.
+To optimize the $\alpha$ values, we'll follow the model fitting procedure up to the point in which the paramter is used, estimate the model parameter using a grid of possible values for the turning parameter, and then find which value for the tuning parameter yields the model parameter estimate which has the lowest root mean square error (RMSE).
+RMSE is calculated by finding the difference between the rating and the rating that is estimated using only the global effects parameter(s).
 
+The first $\alpha$ to be tuned is $\alpha_{game}$.
+The data is randomly divided into 10 separate sections of roughly equal size.
+Then for each of the 10 sections, we'll treat the section as a test set and the other 9 to use as a training set.
+$\widehat{\theta}_g$ is fit first, and then from a grid of $\alpha_{game}$ values (where $\alpha_{game} \in [1,10000]$), a value of $\widehat{\theta}_i$ is produced for each of the $\alpha_{game}$ in the grid.
+For each of the $\widehat{\theta}_i$ that are estimated, we'll use the section that wasn't in the training set, the RMSE is calculated by:
+\begin{equation}
+\text{RMSE} = \frac{ \sum r_{u,i}^{(1)} - (\theta_g + \theta_i)}{n}
+\end{equation}
+
+where we sum over all rated games in the test set, and $n$ is the number of ratings in the test set.
+The value of $\alpha_{game}$ that has the lowest RMSE will be used.
+in figure ~\ref{fig:gamealpha}, we see the optimal value for $\alpha_{game}$ sits in a nice trough, with the minimum at $\alpha_{game} = 1.802303$.
+
+\begin{center}
 \begin{figure}
-<<game_alpha_optim, echo = F>>=
+<<game_alpha_optim, echo = F, fig.width = 4, fig.height = 4 >>=
 load("game_graph")
@@
 @
-\caption{\label{fig:alpha} You need a caption here - what is shown?}
+\caption{\label{fig:gamealpha} RMSE of candidate $\alpha_{game}$ values.}
 \end{figure}
+\end{center}
 
-<<player_alpha_optim, echo = F>>=
+After finding the optimal value for $\alpha_{game}$ we can move on to optimizing $\alpha_{player}$.
+Once again, the data will be divided randomly into approximately 10 equally sized sections, and each will be treated as an individual test set while the other 9 combined will serve as a training set.
+A grid of $\alpha_{player} \in  [1,10000]$ will be used to create a value for $\widehat{\theta}_{u}$, and for each of those values, we'll calculate the RMSE as below:
+
+\begin{equation}
+\text{RMSE} = \frac{ \sum r_{u,i}^{(1)} - (\theta_g + \theta_i + \theta_u)}{n}
+\end{equation}
+
+Once again, the value of $\alpha_{player}$ with the lowest RMSE will be the value used, which as can be seen in ~\ref{fig:playeralpha} is $\alpha_{player} = 2.022021$.
+\begin{figure}
+<<player_alpha_optim, echo = F, fig.width = 4, fig.height = 4>>=
 load("player_graph")
@@
 @
+\caption{\label{fig:playeralpha} RMSE of candidate $\alpha_{player}$ values.}
+\end{figure}
 
+One may immediately notice that different $\alpha$ candidates have a larger effect on RMSE for players than they have for games.
+The reason for this is that there are far more ratings for each game than there are for each player.
+This results in less overfitting for each game and more overfitting for each player, as is seen in the histograms for the number of ratings for each game and player.
+
+\begin{figure}
+<<hists, echo = F, fig.width = 4, fig.height = 4>>=
+load("game_n_rats")
+load("plyr_n_rats")
+game_n_plot
+plyr_n_plot
+@
+\caption{\label{fig:ratn} Frequency of number of ratings for players and games.}
+\end{figure}
+
+\begin{figure}
 <<k_optim, echo = F>>=
@@
 @
-
-While the approaches for finding the optimal values for these parameters may be slightly different, the general implementation will be the same.
-We'll first begin by partitioning a subset of the data into 10 different sections (10 'folds'), and then choosing an appropriate parameter space for the parameter to be optimized.
-Ffor the alpha's, we'll choose values between 0 and 1000 (double check this), evenly spaced, and for $k$ we'll choose a positive integer.
-We'll then proceed fold by fold, using the other 9 folds to fit a model, and then using the data in current fold to check how the model works.
-Using root mean square error, we can then rank each of the possible parameter values in the parameter space, and the value with the lowest rmse will be the 'optimal' value for that parameter.
+\caption{\label{fig:k}}
+\end{figure}
 

### 2015-03-26 paper.Rnw
just a few comments

@@
 \hline
-Family &  Some games a naturally part of a "family" \\
+Family &  Some games a naturally part of a ``family" \\
 &  that share a name or common element \\
@@
 To keep the information as compact as possible, individual players and games are referred to by an ID number, which corresponds to a row in one of the key tables.
-These key tables contain information about the players and game that we don't necessarily need to include with each rating or classifier.
+These key tables contain information about the players and game that we do not necessarily need to include with each rating or classifier.
 This greatly reduces the size of the database needed to store and work with the information.
@@
 Noting this negligible increase in accuracy and recognizing my lack of significant computing resources, we will only use user and game effects.
+\hh{XXX What is your computing power?
+How much further could you go with (how much?) more?
+-- How much more accuracy would you hope for with more power?
+}
 
@@
 Then, you could find people who rated games similarly to you, and use their ratings of the new game as an estimate.
-This often doesnt work, as sometimes the number of people who have rated the same games as you can be relatively low, if not 0 (show a graph for this part).
+This often does not work, as sometimes the number of people who have rated the same games as you can be relatively low, if not 0 (show a graph for this part).
 
@@
 
-\subsection{Model Structure}
+\subsection{Model \hh{Sequence}}
 The specifics of the model follow very closely to those outlined in the Global Effects and knn sections of the BigChaos paper [need bib]. Here, we cover the specifics of both.
@@
 Note for editing: I know this model is a mess at the moment, just getting the general ideas down on paper.
-\hh{this is not nearly as messy as you make it out to be, but you need to make sure to define every little piece of notation as you, because otherwise your reader will be lost.}
+
+\hh{XXX Describe the sequential nature of the models in the next subsections here.}
+
 \subsubsection{Global Effects}
@@
 
-\hh{XXX: generally:  no future tense in the write up. The CC is in the here and now :)
-}
-
-\hh{XXX it is $\theta_u$, not $\theta$. }
-\hh{XXX Why do you call $\alpha$ a fitting parameter and not a shrinking parameter?}
-
 Noting that this can produce esimates that are above 10 and below 1, ratings that fall outside of $[1,10]$ are truncated to the closest interval endpoint.
@@
 All in all, there are up to 1,851 different quantifiable ways that were chosed to descibe games.
-For example, the game Dragonmaster was published in 1981, has a playingtime of 30 minutes, is under the family "Animals: Dragons", falls under the categories "Card Game", "Fantasy Game", is in the subdomain "Strategy Games", uses the mechanic "Trick-taking", and seats 3 or 4 players.
+For example, the game Dragonmaster was published in 1981, has a playingtime of 30 minutes, is under the family ``Animals: Dragons", falls under the categories ``Card Game", ``Fantasy Game", is in the subdomain ``Strategy Games", uses the mechanic ``Trick-taking", and seats 3 or 4 players.
 Each one of these will be treated as a classifier for the game.
 
-In order to model the remaining residuals, we'll give each player a score based on how they rated games with similar classifiers; that is, how did they rate games with the "Trick-taking" mechanic, or how did they rate "Card Games"?
+In order to model the remaining residuals, we'll give each player a score based on how they rated games with similar classifiers; that is, how did they rate games with the ``Trick-taking" mechanic, or how did they rate ``Card Games"?
 
@@
 where the sum is over all residuals that belong to a game that can be described by classifier $z$.
-For example, if the classifier was the mechanic "Trick-taking" and the user was $u = 3$, the classifier score would be the mean of all of the $\epsilon_{3,i}^{(3)}$ for \textit{all} games that the user has rated that have the "Trick-taking" mechanic.
+For example, if the classifier was the mechanic ``Trick-taking" and the user was $u = 3$, the classifier score would be the mean of all of the $\epsilon_{3,i}^{(3)}$ for \textit{all} games that the user has rated that have the ``Trick-taking" mechanic.
 If its the case that a user has rated no games with a particular classifier, we'll leave it as an unobserved value.
@@
 For each pair of players, the correlation between them can be found by finding all of the classifiers that have scores for both players, and then finding the correlation between those values.
-To make sure that we don't find high correlations between players who only have very few classifier scores in common, we'll on consider a correlation between players if they have more than 6 common classifier scores.
+To make sure that we do not find high correlations between players who only have very few classifier scores in common, we'll on consider a correlation between players if they have more than 6 common classifier scores.
 
@@
 
-\subsubsection{Optimizing $\alpha$}
-To optimize the $\alpha$ values, we'll follow the model fitting procedure up to the point in which the paramter is used, estimate the model parameter using a grid of possible values for the turning parameter, and then find which value for the tuning parameter yields the model parameter estimate which has the lowest root mean square error (RMSE).
+\subsubsection{Optimizing tuning parameters $\alpha$}
+\hh{XXX avoid using symbols without also calling them by their names ... i.e. say tuning parameter $\alpha$ rather than just $\alpha$.}
+
+To optimize the $\alpha$ values, we follow the model fitting procedure up to the point in which the paramter is used, estimate the model parameter using a grid of possible values for the turning parameter, and then find which value for the tuning parameter yields the model parameter estimate which has the lowest root mean square error (RMSE).
+\hh{XXX back this definition of RMSE up with a formula and move it to the first time where you mention RMSE. }
 RMSE is calculated by finding the difference between the rating and the rating that is estimated using only the global effects parameter(s).
 
+\hh{What does 'the first'meane here?
+Does it mean that there is a given order, which we have to follow, or did you decide to first tune the game parameter?
+Does it matter which parameter is tuned first?}
 The first $\alpha$ to be tuned is $\alpha_{game}$.
+\hh{XXX What you are describing here is called ten-fold cross validation. }
 The data is randomly divided into 10 separate sections of roughly equal size.
@@
 in figure ~\ref{fig:gamealpha}, we see the optimal value for $\alpha_{game}$ sits in a nice trough, with the minimum at $\alpha_{game} = 1.802303$.
+\hh{XXX do you really need that many significant digits?}
 
@@
 \begin{figure}
-<<game_alpha_optim, echo = F, fig.width = 4, fig.height = 4 >>=
+\centering
+<<game_alpha_optim, echo = F, fig.width = 4, fig.height = 4, out.width='0.5\\textwidth'>>=
 load("game_graph")
@@
 @
-\caption{\label{fig:gamealpha} RMSE of candidate $\alpha_{game}$ values.}
+\caption{\label{fig:gamealpha} RMSE of candidate $\alpha_{game}$ values. \hh{XXX there's a typo in the x label }}
 \end{figure}
@@
 \begin{figure}
-<<player_alpha_optim, echo = F, fig.width = 4, fig.height = 4>>=
+\centering
+<<player_alpha_optim, echo = F, fig.width = 4, fig.height = 4, out.width='0.5\\textwidth'>>=
 load("player_graph")
@@
 @
-\caption{\label{fig:playeralpha} RMSE of candidate $\alpha_{player}$ values.}
+\caption{\label{fig:playeralpha} RMSE of candidate $\alpha_{player}$ values. \hh{typo in X}}
 \end{figure}
@@
 \begin{figure}
-<<hists, echo = F, fig.width = 4, fig.height = 4>>=
+\centering
+<<hists, echo = F, fig.width = 4, fig.height = 4, out.width='0.45\\textwidth', fig.show='hold'>>=
 load("game_n_rats")
@@
 @
-\caption{\label{fig:ratn} Frequency of number of ratings for players and games.}
+\caption{\label{fig:ratn} Frequency of number of ratings for games (left) amd players (right).
+\hh{A log transformation on the frequencies might be more informative.
+Make sure to emphasize it, if you show transformed data.}}
 \end{figure}
@@
 
-% don't box citations, add space between multiple citations, separate with ;
+% do not box citations, add space between multiple citations, separate with ;
 \def\@citex[#1]#2{\if@filesw\immediate\write\@auxout{\string\citation{#2}}\fi

### 2015-03-31 paper.Rnw
Wrapped up
Wrapped up Tech. Implementation and Conclusion and responded to
comments. Included screenshots and additional graphs.

@@
 \usepackage[normalem]{ulem}
+\usepackage{graphicx}
 
@@
 \begin{abstract}
-\hh{You'll need to include an abstract here}
+This project implements portions of Netflix-Prize finalist's algorithms in an attempt to predict how much a person will enjoy board games.
+Drawing on user's past predictions and data on the specifics of each game, a heirarchical method is used to link a specific user's ratings to ratings from other users, and allowing the use of data on other users who may have not rated any of the same games to predict enjoyment of unrated games.
+Through this model, we also can derive user-specific scores on how much they like certain aspects of games.
 \end{abstract}
@@
 %the small desktop I have to work on.
-\hh{Instead, we} %Secondly, this will force us to 
-implement a continuous data scraping procedure \hh{and update the recommender model dynamically:} Each time a person uses the app, %all we have to do is 
+Instead, we %Secondly, this will force us to 
+implement a continuous data scraping procedure and update the recommender model dynamically: Each time a person uses the app, %all we have to do is 
 we make sure to check that all of their current information is correct.
@@
 \subsection{Model Selection}
-\hh{We base our implementation of a recommender system} %To begin implementing such a model, we can look to 
-on two papers  of on one the most famous recommender systems of all: Netflix.
+We base our implementation of a recommender system %To begin implementing such a model, we can look to 
+on two papers which aimed to improve upon one the most famous recommender systems of all: Netflix.
 In 2006, Netflix offered a \$1 Million prize for an improvement on their algorithm for predicting user rating on movies they had not seen.
@@
 
-\hh{XXX What is the approach you are taking compared to Netflix? the same? similar? what is different? - One or two sentences here.}
+Our approach is to borrow from a single element of their shotgun modeling. Where they use many different models of many different types, we will choose a simple implementation of a single model.
 
@@
 We will implement these in a similar manner to the way that BigChaos implemented the Global Effects model \citep{bigchaos}.
-\hh{XXX the next clause needs an end to be a sentence :)}
-While they implemented 14 steps including Movie Effect, User Effect, and those variables crossed with variables like time since rating, average rating for the movie, production year, standard deviations, as well as others.
-\hh{XXX explain RMSE; make sure in the next paragraph to always point out what is yours and what is netflix. You know all those distinctions, but for any other reader it is not that obvious.}
-From the RMSE of their probe set, the difference between using only user/movie effcts and using all 14 effects was only .02, a decrease of about 2\%.
+While they implemented 14 steps including Movie Effect, User Effect, and those variables crossed with variables like time since rating, average rating for the movie, production year, standard deviations, as well as others, we'll stick to only using 2 steps: an effect for each game and an effect for each player.
+To compare models to each other we use root mean square error (RMSE).
+We take our data and split it into a training and test set.
+The model is fit with the training data, and then we predict the values in the test set.
+From those values, we can then find the RMSE using the equation below, where $y$ are the observed values from the test set, $\widehat{y}$ are the predicted values from the model built with the training set, and $n$ is the number of observations in the test dataset.
+
+\begin{equation}
+\text{RMSE} = \sqrt{\frac{\sum (y - \widehat{y})^2}{n}}
+\end{equation}
+
+It is important to note that RMSE is a relative measure, and unless you're comparing models on the same data, a similar magnitude of change in RMSE between two different datasets can be meaningless.
+Therefore when we look at the RMSE results on the Netflix data, we look to the RMSE change as guidance for our model knowing that we may see different RMSE changes when we implement it with different data.
+% \hh{XXX explain RMSE; make sure in the next paragraph to always point out what is yours and what is netflix. You know all those distinctions, but for any other reader it is not that obvious.}
+From the RMSE of their probe set of the Netflix data, the difference between using only user/movie effcts and using all 14 effects was only .02, a decrease of about 2\%.
 Noting this negligible increase in accuracy and recognizing my lack of significant computing resources, we will only use user and game effects.
-\hh{XXX What is your computing power?
-How much further could you go with (how much?) more?
--- How much more accuracy would you hope for with more power?
-}
+My model is fit on my home desktop, which has an Intel Core i7-4770K CPU @ 3.5GHz and 32GB of RAM running 64-bit Windows 8.
+With the help of additional computing cores, RAM, and better parallelization, we could hope to acheive the 2\% accuracy increase seen in the Netflix Data, but more than likely it wouldn't be worth it, as a quick cost/benefit analysis reveals that I don't need to hit the nail on the head every time, I just need to implement a method that reliably works a good enough percentage of the time.
+To this coder, a 2\% increase in accuracy just isn't worth the money.
 
@@
 
-\subsection{Model \hh{Sequence}}
-The specifics of the model follow very closely to those outlined in the Global Effects and knn sections of the BigChaos paper [need bib]. Here, we cover the specifics of both.
-
-Note for editing: I know this model is a mess at the moment, just getting the general ideas down on paper.
-
-\hh{XXX Describe the sequential nature of the models in the next subsections here.}
+\subsection{Model Sequence}
+The specifics of the model follow very closely to those outlined in the Global Effects and knn sections of the BigChaos paper [need bib].
+Here, we cover the specifics of both.
+We will step through the model in sequence, removing the effects that games and players natrually place on ratings (some games are naturally rated higher than others, some players are more or less generous than others).
+After removing those larger effects, we'll model the remaining nuance using charachteristics about the games that were rated.
+From there, we'll strive to find users who have similar tastes in game characteristics, and then finally use that closeness to find ratings for games that a user has not yet rated.
 
@@
 Each one of these will be treated as a classifier for the game.
+A table storing this information by GAME\_ID can be found in Table 2.
 
@@
 \subsubsection{Optimizing tuning parameters $\alpha$}
-\hh{XXX avoid using symbols without also calling them by their names ... i.e. say tuning parameter $\alpha$ rather than just $\alpha$.}
 
-To optimize the $\alpha$ values, we follow the model fitting procedure up to the point in which the paramter is used, estimate the model parameter using a grid of possible values for the turning parameter, and then find which value for the tuning parameter yields the model parameter estimate which has the lowest root mean square error (RMSE).
-\hh{XXX back this definition of RMSE up with a formula and move it to the first time where you mention RMSE. }
+To optimize the tuning parameter $\alpha$ values, we follow the model fitting procedure up to the point in which the paramter is used, estimate the model parameter using a grid of possible values for the turning parameter, and then find which value for the tuning parameter yields the model parameter estimate which has the lowest RMSE (equation (1)).
 RMSE is calculated by finding the difference between the rating and the rating that is estimated using only the global effects parameter(s).
+Since the game parameter is fit first, we fit the $\alpha$ tuning parameter for the game effect first.
+After finding an optimal value, we'll then fit the $\alpha$ tuning parameter for the player effect, as the game effect is needed to estimate the player effect.
 
-\hh{What does 'the first'meane here?
-Does it mean that there is a given order, which we have to follow, or did you decide to first tune the game parameter?
-Does it matter which parameter is tuned first?}
-The first $\alpha$ to be tuned is $\alpha_{game}$.
-\hh{XXX What you are describing here is called ten-fold cross validation. }
-The data is randomly divided into 10 separate sections of roughly equal size.
+The optimization will use ten-fold cross validation, wherein the data is randomly divided into 10 separate sections of roughly equal size.
 Then for each of the 10 sections, we'll treat the section as a test set and the other 9 to use as a training set.
@@
 The value of $\alpha_{game}$ that has the lowest RMSE will be used.
-in figure ~\ref{fig:gamealpha}, we see the optimal value for $\alpha_{game}$ sits in a nice trough, with the minimum at $\alpha_{game} = 1.802303$.
-\hh{XXX do you really need that many significant digits?}
+in figure ~\ref{fig:gamealpha}, we see the optimal value for $\alpha_{game}$ sits in a nice trough, with the minimum at $\alpha_{game} = 1.802$.
 
@@
 <<game_alpha_optim, echo = F, fig.width = 4, fig.height = 4, out.width='0.5\\textwidth'>>=
-load("game_graph")
+load("game_graph2")
 game_graph
 @
-\caption{\label{fig:gamealpha} RMSE of candidate $\alpha_{game}$ values. \hh{XXX there's a typo in the x label }}
+\caption{\label{fig:gamealpha} RMSE of candidate $\alpha_{game}$ values.}
 \end{figure}
@@
 
-Once again, the value of $\alpha_{player}$ with the lowest RMSE will be the value used, which as can be seen in ~\ref{fig:playeralpha} is $\alpha_{player} = 2.022021$.
+Once again, the value of $\alpha_{player}$ with the lowest RMSE will be the value used, which as can be seen in ~\ref{fig:playeralpha} is $\alpha_{player} = 2.022$.
 \begin{figure}
@@
 <<player_alpha_optim, echo = F, fig.width = 4, fig.height = 4, out.width='0.5\\textwidth'>>=
-load("player_graph")
+load("player_graph2")
 player_graph
 @
-\caption{\label{fig:playeralpha} RMSE of candidate $\alpha_{player}$ values. \hh{typo in X}}
+\caption{\label{fig:playeralpha} RMSE of candidate $\alpha_{player}$ values.}
 \end{figure}
@@
 The reason for this is that there are far more ratings for each game than there are for each player.
-This results in less overfitting for each game and more overfitting for each player, as is seen in the histograms for the number of ratings for each game and player.
+This results in less overfitting for each game and more overfitting for each player, as is seen in the histograms for the log number of ratings for each game and player.
+Log was used as the number of ratings for player and game is highly right skewed.
 
@@
 <<hists, echo = F, fig.width = 4, fig.height = 4, out.width='0.45\\textwidth', fig.show='hold'>>=
-load("game_n_rats")
-load("plyr_n_rats")
+load("game_n_rats2")
+load("plyr_n_rats2")
 game_n_plot
@@
 @
-\caption{\label{fig:ratn} Frequency of number of ratings for games (left) amd players (right).
-\hh{A log transformation on the frequencies might be more informative.
-Make sure to emphasize it, if you show transformed data.}}
+\caption{\label{fig:ratn} Frequency of log number of ratings for games (left) amd players (right).}
 \end{figure}
@@
 \section{Technical Implementation}
-** SCREENSHOTS NEEDED **
-
 This algorithm is implemented in R using shiny as a vehhicle for a simple, easy to use user interface.
@@
 In an implementation of this, the application will begin by having the user enter the username they've registered on \url{boardgamegeek.com}.
+As seen in (~\ref{fig:blanksc}), there is an area to enter the users information in, and a button just below that, when clicked, begins the process.
+\begin{center}
+\begin{figure}
+\includegraphics[scale=0.4]{blank_login.png}
+\caption{\label{fig:blanksc} A blank login screen}
+\end{figure}
+\end{center}
 Instead of logging the users ratings directly through the application, we will instead leverage the existing infrastructure of boardgame geek.
@@
 Have they rated games that aren't in our database?
-.If so, we also need to bring in other users ratings for any games that we may not have had in our database (since a single rating for a game isn't very useful and will greatly skew the results for that game).
+If so, we also need to bring in other users ratings for any games that we may not have had in our database (since a single rating for a game isn't very useful and will greatly skew the results for that game).
 After this initial step and database update, we can see if they already have saved predicted ratings in the system, and if not, we'll proceed to model fitting.
 
-We first apply the global effects strategy as descibed in the modling section.
-The optimization of the $\alpha$ parameters will have occred sometime before this and the algoritm will use use those stored values.
-A weeky or monthly automatic updaing procedure is used to keep these values up to date as more data comes into the system.
+We first apply the global effects strategy as descibed in the modeling section.
+The optimization of the $\alpha$ tuning parameters will have occred sometime before this and the algoritm will use use those stored values.
+A weeky or monthly automatic updaing procedure is to be used to keep these values up to date as more data comes into the system.
 From the global effects, we then aggregate those based on the classifier aggregation procedure as described above.
@@
 
-Finally, after the k-fold crossvalidation, we're ready to run the other games through the algorithm to get a predicted score.
-The games that the user has already rated will then be replaced with those ratings, and the full list will be outputted for the user.
+Finally, after the 10-fold cross-validation, we're ready to run the other games through the algorithm to get a predicted score.
+The games that the user has already rated will then be replaced with those ratings, and the full list will be outputted for the user in the "My Recommendations" Tab, which can be seen in Figure ~\ref{fig:act_use}.
+\begin{figure}
+\includegraphics[scale=0.4]{active_use.png}
+\caption{\label{fig:act_use} The "My Recommendations" tab, populated for user cwmassey.}
+\end{figure}
 
 Since we saved the aggregated classifier ratings for the individual user, we can suggest ways for the user to subset this data.
-By providing the graph below to the player (GET THE GRAPH WORKING), they can see for themselves their not-necessarily obvious preferences in game family or mechanic.
-They can then search by their preferred mechanic or theme, and get ratings for games that fit that description.
-Or the user can simply sort by highest rated.
+By providing the graph in Figure ~\ref{fig:subdplot} below to the player they can see for themselves their not-necessarily obvious preferences in game family or mechanic (in this case Subdomain).
+For this particular user, this would indicate them that they tend to prefer Customizable or Party games rather than Children's Games or Strategy Games.
+They can then search by their preferred subdomain, and get ratings for games that fit that description.
+This plot is provided to the user within the "My Category Ratings" tab.
+The plot for the specific category rating can be chosen by the user, so the user can select from Category, Family, Mechanic, Subdomain, Number of Players, Playing Time, or Year Published.
+
+\begin{figure}
+<<subdom_plot, echo = FALSE, fig.width = 4, fig.height = 4, out.width='0.5\\textwidth'>>=
+load("classifiers_plot")
+out
+@
+\caption{\label{fig:subdplot} A player preference plot for subdomains for the user cwmassey.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.4]{cat_ratings.png}
+\caption{\label{fig:cat_rat_tab} The "My Category Ratings" tab.}
+\end{figure}
+
+If a user wants to rate a new game, they have to go back to BoardGameGeek and rate games there. The app will then need to be restarted, which will pull the new rating, and update the database.
 
 \section{Conclusion}
+While the result was not as smooth as hoped, the Netflix recommender system made an excellent base off of which to build a recommender system for board games.
+It would not be too much of a stretch to imagine that, if you had a reasonable amount of user rating data about something, and you had enough descriptive data about that thing, that a similar implementation would be of great use there as well.
+
+There are a few directions that this project is naturally headed in.
+The first of which is to speed up the prediction generating process.
+Currently, each time a user starts the app with new data, the new data is fetched and the whole model is fit again, including generating a truly massive correlation matrix.
+Instead of doing this each time a person logs in, it may be more advantageous to fit the correlation matrix daily, and then only update the global effects portion of the model for each user, as that portion is relatively snappy.
+Hopefully this would bring the fitting time down from 20 minutes to a matter of seconds.
+Additionally, we don't have the full data that is contained on BoardGameGeek.
+There is much more data to scrape than time for this paper allowed, so going and continually updating the database would be advantageous.
+
+Another possible step is to generalize the whole recommendation system procedure, so that if you had rating and classifing information on any manner of subjects, you could implement a recommendation system with little to no effort.
+This could have a whole host of applications from what restaurants you might enjoy to what books you might like to read.
+
+Finally, there are hosts of other recommender systems out there.
+Augmenting this approach with additional models as done in the Netflix Prize papers would undoubably be a logical step forward for this kind of modeling.
 

### 2015-03-31 paper.Rnw
RMSE definition, some typos

@@
 This project implements portions of Netflix-Prize finalist's algorithms in an attempt to predict how much a person will enjoy board games.
-Drawing on user's past predictions and data on the specifics of each game, a heirarchical method is used to link a specific user's ratings to ratings from other users, and allowing the use of data on other users who may have not rated any of the same games to predict enjoyment of unrated games.
+Drawing on user's past predictions and data on the specifics of each game, a hierarchical method is used to link a specific user's ratings to ratings from other users, and allowing the use of data on other users who may have not rated any of the same games to predict enjoyment of unrated games.
 Through this model, we also can derive user-specific scores on how much they like certain aspects of games.
@@
 A pretty thorough analysis of it (and its issues) can be found at \url{http://boardgamegeek.com/wiki/page/Game_Recommendation_Algorithm}.
-The bigget issue is that it cannot be tailored to an individual.
-You, an individual user, may have different preferences than the userbase average, and it might be more useful to tailor recommendations to the individual.
+The bigger issue is that it cannot be tailored to an individual.
+You, an individual user, may have different preferences than the user base average, and it might be more useful to tailor recommendations to the individual.
 
-Recommendations for the individual aren't easy to come by.
-There are a few automated systems out there (such as the boargamerecommender bot on reddit), but for the most part they are time intensive and often lack interactivity or any sort of information that would be useful to the user beyond a few single games that it might recommend.
+Recommendations for the individual aren't easy to come by. There are a few automated systems out there (such as the boardgame 
+recommender bot on reddit), but for the most part they are time intensive and often lack interactivity or any sort of information that would be useful to the user beyond a few single games that it might recommend.
 
@@
 Users can create an account and then rate games they've played as well as keep an inventory of games they own.
-All of this information is accessable and scrapable, and will be the data that drives our analysis.
+All of this information is accessible and scrape-able, and will be the data that drives our analysis.
 
@@
 
-Not all of the data was scraped before creating the applicaiton.
+Not all of the data was scraped before creating the application.
 Data for 6,088 games was scraped and stored across the four tables.
@@
 First of all, for those 6,088 games there are over 3 million unique ratings.
-If we were to expand it to all 75,756 games known to BoardGameGeek, then our data set would be crippingly large for all but very high-powered machines.
+If we were to expand it to all 75,756 games known to BoardGameGeek, then our data set would be cripplingly large for all but very high-powered machines.
 %the small desktop I have to work on.
@@
 \begin{equation}
-\text{RMSE} = \sqrt{\frac{\sum (y - \widehat{y})^2}{n}}
+\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^n \left(y_i - \widehat{y_i}\right)^2
 \end{equation}
 
-It is important to note that RMSE is a relative measure, and unless you're comparing models on the same data, a similar magnitude of change in RMSE between two different datasets can be meaningless.
+It is important to note that RMSE is a relative measure, and unless we are comparing models on the same data and subset, a similar magnitude of change in RMSE between two different datasets can be meaningless.
 Therefore when we look at the RMSE results on the Netflix data, we look to the RMSE change as guidance for our model knowing that we may see different RMSE changes when we implement it with different data.
 % \hh{XXX explain RMSE; make sure in the next paragraph to always point out what is yours and what is netflix. You know all those distinctions, but for any other reader it is not that obvious.}
-From the RMSE of their probe set of the Netflix data, the difference between using only user/movie effcts and using all 14 effects was only .02, a decrease of about 2\%.
+From the RMSE of the probe set of the Netflix data, the difference between using only user/movie effects and using all 14 effects was .02, a decrease of about 2\%.
 Noting this negligible increase in accuracy and recognizing my lack of significant computing resources, we will only use user and game effects.
 My model is fit on my home desktop, which has an Intel Core i7-4770K CPU @ 3.5GHz and 32GB of RAM running 64-bit Windows 8.
-With the help of additional computing cores, RAM, and better parallelization, we could hope to acheive the 2\% accuracy increase seen in the Netflix Data, but more than likely it wouldn't be worth it, as a quick cost/benefit analysis reveals that I don't need to hit the nail on the head every time, I just need to implement a method that reliably works a good enough percentage of the time.
+With the help of additional computing cores, RAM, and better parallelization, we could hope to achieve the 2\% accuracy increase seen in the Netflix Data, but more than likely it wouldn't be worth it, as a quick cost/benefit analysis reveals that I don't need to hit the nail on the head every time, I just need to implement a method that reliably works a good enough percentage of the time.
 To this coder, a 2\% increase in accuracy just isn't worth the money.
@@
 After we fit the general effects, we wish to model the remaining residuals, ideally in some way that gives us informative numeric summaries about the individual users in the process.
-In addition, we'd also like to compare a users ratings to other users, and hopefully use the behavior of simiar users to predict new ratings for someone.
+In addition, we'd also like to compare a users ratings to other users, and hopefully use the behavior of similar users to predict new ratings for someone.
 A tempting first step is to, for a particular game, find other people who have rated that game who have rated games that you've rated.
@@
 Here, we cover the specifics of both.
-We will step through the model in sequence, removing the effects that games and players natrually place on ratings (some games are naturally rated higher than others, some players are more or less generous than others).
-After removing those larger effects, we'll model the remaining nuance using charachteristics about the games that were rated.
+We will step through the model in sequence, removing the effects that games and players naturally place on ratings (some games are naturally rated higher than others, some players are more or less generous than others).
+After removing those larger effects, we'll model the remaining nuance using characteristics about the games that were rated.
 From there, we'll strive to find users who have similar tastes in game characteristics, and then finally use that closeness to find ratings for games that a user has not yet rated.
@@
 
-We'll estimate these effects one at a time using a heirarchical structure.
+We'll estimate these effects one at a time using a hierarchical structure.
 Define $r_{u,i}^{(1)}$ be the rating from user $u$ for game $i$.
@@
 For this purpose, define $r_{u,i}^{(2)} = \epsilon_{u,i}^{(1)}$, which we'll model using global effect for each game, denoted $\theta_i$.
-The residuals are then modeled using the following, where $x_i$ is an indicator varialbe that the residual belongs to game $i$:
+The residuals are then modeled using the following, where $x_i$ is an indicator variable that the residual belongs to game $i$:
 
@@
 The final step in the global effects stage is finding the global effect for each user denoted $\theta_{u}$.
-In a simillar manner to the game global effects, once again taking the residual from the previous step and model them.
+In a similar manner to the game global effects, once again taking the residual from the previous step and model them.
 To proceed, let $r_{u,i}^{(3)} = \epsilon_{u,i}^{(2)}$, and those residuals are modeled using:
@@
 
-Noting that this can produce esimates that are above 10 and below 1, ratings that fall outside of $[1,10]$ are truncated to the closest interval endpoint.
+Noting that this can produce estimates that are above 10 and below 1, ratings that fall outside of $[1,10]$ are truncated to the closest interval endpoint.
 %  The general idea is to remove the global mean from all ratings to get a residual for each observation.
@@
 Each game in the dataset can be described by things like what mechanics it has, or what family of game it falls under, when it was published, etc.
-All in all, there are up to 1,851 different quantifiable ways that were chosed to descibe games.
-For example, the game Dragonmaster was published in 1981, has a playingtime of 30 minutes, is under the family ``Animals: Dragons", falls under the categories ``Card Game", ``Fantasy Game", is in the subdomain ``Strategy Games", uses the mechanic ``Trick-taking", and seats 3 or 4 players.
+All in all, there are up to 1,851 different quantifiable ways that were chose to describe games.
+For example, the game Dragonmaster was published in 1981, has a playing time of 30 minutes, is under the family ``Animals: Dragons", falls under the categories ``Card Game", ``Fantasy Game", is in the subdomain ``Strategy Games", uses the mechanic ``Trick-taking", and seats 3 or 4 players.
 Each one of these will be treated as a classifier for the game.
@@
 
-To optimize the tuning parameter $\alpha$ values, we follow the model fitting procedure up to the point in which the paramter is used, estimate the model parameter using a grid of possible values for the turning parameter, and then find which value for the tuning parameter yields the model parameter estimate which has the lowest RMSE (equation (1)).
+To optimize the tuning parameter $\alpha$ values, we follow the model fitting procedure up to the point in which the parameter is used, estimate the model parameter using a grid of possible values for the turning parameter, and then find which value for the tuning parameter yields the model parameter estimate which has the lowest RMSE (equation (1)).
 RMSE is calculated by finding the difference between the rating and the rating that is estimated using only the global effects parameter(s).
@@
 @
-\caption{\label{fig:ratn} Frequency of log number of ratings for games (left) amd players (right).}
+\caption{\label{fig:ratn} Frequency of log number of ratings for games (left) and players (right).}
 \end{figure}
@@
 \section{Technical Implementation}
-This algorithm is implemented in R using shiny as a vehhicle for a simple, easy to use user interface.
-Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that will be useful in the classifier aggregation and k-nearest-neighbors section; the predicted ratings table, which will store predicted ratigs if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user.
-This data storage will allow for ease of implementation: we only really need the ratings table and the classifier table to do all the work, and the key tables are really only needed to display the inforation to the user in a format which they will understand.
+This algorithm is implemented in R using shiny as a vehicle for a simple, easy to use user interface.
+Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that will be useful in the classifier aggregation and k-nearest-neighbors section; the predicted ratings table, which will store predicted ratings if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user.
+This data storage will allow for ease of implementation: we only really need the ratings table and the classifier table to do all the work, and the key tables are really only needed to display the information to the user in a format which they will understand.
 
@@
 \includegraphics[scale=0.4]{blank_login.png}
-\caption{\label{fig:blanksc} A blank login screen}
+\caption{\label{fig:blanksc} A blank log in screen}
 \end{figure}
@@
 Have they rated games that aren't in our database?
-Have they rerated games that are in our database?
+Have they re rated games that are in our database?
 Have they rated games that aren't in our database?
@@
 
-We first apply the global effects strategy as descibed in the modeling section.
-The optimization of the $\alpha$ tuning parameters will have occred sometime before this and the algoritm will use use those stored values.
-A weeky or monthly automatic updaing procedure is to be used to keep these values up to date as more data comes into the system.
+We first apply the global effects strategy as described in the modeling section.
+The optimization of the $\alpha$ tuning parameters will have occurred sometime before this and the algorithm will use use those stored values.
+A weekly or monthly automatic updating procedure is to be used to keep these values up to date as more data comes into the system.
 From the global effects, we then aggregate those based on the classifier aggregation procedure as described above.
@@
 
-Another possible step is to generalize the whole recommendation system procedure, so that if you had rating and classifing information on any manner of subjects, you could implement a recommendation system with little to no effort.
+Another possible step is to generalize the whole recommendation system procedure, so that if you had rating and classifying information on any manner of subjects, you could implement a recommendation system with little to no effort.
 This could have a whole host of applications from what restaurants you might enjoy to what books you might like to read.
@@
 Finally, there are hosts of other recommender systems out there.
-Augmenting this approach with additional models as done in the Netflix Prize papers would undoubably be a logical step forward for this kind of modeling.
+Augmenting this approach with additional models as done in the Netflix Prize papers would undoubtedly be a logical step forward for this kind of modeling.
 

### 2015-03-31 paper.Rnw
couple of comments

@@
 
-\begin{equation}
-\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^n \left(y_i - \widehat{y_i}\right)^2
+\begin{equation}\label{eq:rmse}
+\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^n \left(y_i - \widehat{y_i}\right)^2}
 \end{equation}
@@
 
-Define $C_{u,z}$ to the be the classifier score for user $u$ for classifier $z$ where $z$ is an index of the 1,851 classifiers. Then:
+Define $C_{u,z}$ to the be the classifier score for user $u$ for classifier $z$ where $z$ is an index of the 1,851 classifiers.
+Then: \hh{XXX something went wrong with the formula ...
+not sure how to fix it}
 
@@
 
-The weights for this will be $\frac{1}{1 + \rho_{u,k}}$ where $\rho_{u,k}$ is the correlation between user $u$ and user $k$.
+\hh{XXX why do those weights make sense?
+aren't those weights tilted towards players who are negatively correlated?
+are there any?
+(then it's not  a practical problem, even though still a theoretical one) just check that it is not supposed to be $1 - \rho$ in the denominator.
+I'm pretty sure that's what's it supposed to be.}
+The weights for this will be $(1 + \rho_{u,k})^{-1}$ where $\rho_{u,k}$ is the correlation between user $u$ and user $k$.
 The number of neighbors to use $k$ will be determined by cross-validation on a user-to-user basis.
-In the end $z_{u,i}$ will be the k-nearest-neighbors estimate for user $u$ and game $i$, as given in the equation below (where $u_{[j]}$ is the user with the $j^{\text{th}}$ highest correlation between themselves and user $u$:
+In the end $z_{u,i}$ will be the \hh{weighted average of the} k-nearest-neighbors \hh{scores} for user $u$ and game $i$, as given in the equation below (where $u_{[j]}$ is the user with the $j^{\text{th}}$ highest correlation between themselves and user $u$):
 
 \begin{equation}
-z_{u,i} = \frac{\sum_{j = 1}^{k} \epsilon_{u_{[j]},i}^{(3)} * \frac{1}{1 + \rho_{u,u_{[j]}}}}{\sum_{j = 1}^{k}\frac{1}{1 + \rho_{u,u_{[j]}}}}
+z_{u,i} = \frac{\sum_{j = 1}^{k} \epsilon_{u_{[j]},i}^{(3)} \cdot \frac{1}{1 + \rho_{u,u_{[j]}}}}{\sum_{j = 1}^{k}\frac{1}{1 + \rho_{u,u_{[j]}}}}
 \end{equation}
@@
 \subsubsection{Back to Rating}
-From what we've produced above, of the estimated rating for user $u$ and game $i$ is given by:
+From what we've produced above \hh{XXX to be more specific use the relevant equation numbers that are involved }, of the estimated rating for user $u$ and game $i$ is given by:
 
@@
 
-To optimize the tuning parameter $\alpha$ values, we follow the model fitting procedure up to the point in which the parameter is used, estimate the model parameter using a grid of possible values for the turning parameter, and then find which value for the tuning parameter yields the model parameter estimate which has the lowest RMSE (equation (1)).
+To optimize the tuning parameter $\alpha$ values, we follow the model fitting procedure up to the point in which the parameter is used, estimate the model parameter using a grid of possible values for the turning parameter, and then find which value for the tuning parameter yields the model parameter estimate which has the lowest RMSE (equation \ref{eq:rmse}).
 RMSE is calculated by finding the difference between the rating and the rating that is estimated using only the global effects parameter(s).
@@
 $\widehat{\theta}_g$ is fit first, and then from a grid of $\alpha_{game}$ values (where $\alpha_{game} \in [1,10000]$), a value of $\widehat{\theta}_i$ is produced for each of the $\alpha_{game}$ in the grid.
-For each of the $\widehat{\theta}_i$ that are estimated, we'll use the section that wasn't in the training set, the RMSE is calculated by:
+For each of the $\widehat{\theta}_i$ that are estimated, we'll use the section that wasn't in the training set, the RMSE is calculated here by: \hh{XXX I am not sure this really is the RMSE - where did the squaring and the square rooting go?}
 \begin{equation}
@@
 The value of $\alpha_{game}$ that has the lowest RMSE will be used.
-in figure ~\ref{fig:gamealpha}, we see the optimal value for $\alpha_{game}$ sits in a nice trough, with the minimum at $\alpha_{game} = 1.802$.
+in figure~\ref{fig:gamealpha}, we see the optimal value for $\alpha_{game}$ sits in a nice trough, with the minimum at $\alpha_{game} = 1.802$.
 
-\begin{center}
 \begin{figure}
@@
 \end{figure}
-\end{center}
 
@@
 
-\begin{figure}
-<<k_optim, echo = F>>=
+\begin{figure}\centering
+<<k_optim, echo = F, out.width='0.45\\textwidth'>>=
 load("k_plot")
@@
 @
-\caption{\label{fig:k}}
+\caption{\label{fig:k} \hh{You still need a caption here.}}
 \end{figure}
 
-\section{Technical Implementation}
+\section{Technical Implementation} \hh{give references for R (to be a nice R user), shiny, and the other packages that you are using.
+citation() and citation("packagename") makes your life easy with that.}
 This algorithm is implemented in R  using shiny as a vehicle for a simple, easy to use user interface.
@@
 \begin{center}
-\begin{figure}
+\begin{figure}\centering
 \includegraphics[scale=0.4]{blank_login.png}
@@
 The games that the user has already rated will then be replaced with those ratings, and the full list will be outputted for the user in the "My Recommendations" Tab, which can be seen in Figure ~\ref{fig:act_use}.
-\begin{figure}
+\begin{figure}\centering
 \includegraphics[scale=0.4]{active_use.png}
@@
 
-\begin{figure}
+\begin{figure}\centering
 <<subdom_plot, echo = FALSE, fig.width = 4, fig.height = 4, out.width='0.5\\textwidth'>>=
@@
 @
-\caption{\label{fig:subdplot} A player preference plot for subdomains for the user cwmassey.}
+\caption{\label{fig:subdplot} A player preference plot for subdomains for the user cwmassey.  \hh{XXX fix the y label, if showing it}}
 \end{figure}
 
-\begin{figure}
+\begin{figure}\centering
 \includegraphics[scale=0.4]{cat_ratings.png}
-\caption{\label{fig:cat_rat_tab} The "My Category Ratings" tab.}
+\caption{\label{fig:cat_rat_tab} The ``My Category Ratings" tab.}
 \end{figure}

### 2015-04-01 paper.Rnw
Final edits and citations

@@
 
-\hh{Refer to table~\ref{tab:classifiers} somewhere in the text, add a label to the ratings table as well.}
+% \hh{Refer to table~\ref{tab:classifiers} somewhere in the text, add a label to the ratings table as well.}
 <<classifiers, echo = FALSE, results = 'asis'>>=
@@
 
-Define $C_{u,z}$ to the be the classifier score for user $u$ for classifier $z$ where $z$ is an index of the 1,851 classifiers.
-Then: \hh{XXX something went wrong with the formula ...
-not sure how to fix it}
+Define $C_{u,z}$ to the be the classifier score for user $u$ for classifier $z$ where $z$ is an index of the 1,851 classifiers. Then:
 
@@
 
-\hh{XXX why do those weights make sense?
-aren't those weights tilted towards players who are negatively correlated?
-are there any?
-(then it's not  a practical problem, even though still a theoretical one) just check that it is not supposed to be $1 - \rho$ in the denominator.
-I'm pretty sure that's what's it supposed to be.}
-The weights for this will be $(1 + \rho_{u,k})^{-1}$ where $\rho_{u,k}$ is the correlation between user $u$ and user $k$.
+The weights for this will be $(1 - \rho_{u,k})^{-1}$ where $\rho_{u,k}$ is the correlation between user $u$ and user $k$.
 The number of neighbors to use $k$ will be determined by cross-validation on a user-to-user basis.
-In the end $z_{u,i}$ will be the \hh{weighted average of the} k-nearest-neighbors \hh{scores} for user $u$ and game $i$, as given in the equation below (where $u_{[j]}$ is the user with the $j^{\text{th}}$ highest correlation between themselves and user $u$):
+In the end $z_{u,i}$ will be the weighted average of the k-nearest-neighbors scores for user $u$ and game $i$, as given in the equation below (where $u_{[j]}$ is the user with the $j^{\text{th}}$ highest correlation between themselves and user $u$):
 
 \begin{equation}
-z_{u,i} = \frac{\sum_{j = 1}^{k} \epsilon_{u_{[j]},i}^{(3)} \cdot \frac{1}{1 + \rho_{u,u_{[j]}}}}{\sum_{j = 1}^{k}\frac{1}{1 + \rho_{u,u_{[j]}}}}
+z_{u,i} = \frac{\sum_{j = 1}^{k} \epsilon_{u_{[j]},i}^{(3)} \cdot \frac{1}{1 - \rho_{u,u_{[j]}}}}{\sum_{j = 1}^{k}\frac{1}{1 - \rho_{u,u_{[j]}}}}
 \end{equation}
@@
 \subsubsection{Back to Rating}
-From what we've produced above \hh{XXX to be more specific use the relevant equation numbers that are involved }, of the estimated rating for user $u$ and game $i$ is given by:
+From what we've produced above in equations (2)-(9), of the estimated rating for user $u$ and game $i$ is given by:
 
@@
 $\widehat{\theta}_g$ is fit first, and then from a grid of $\alpha_{game}$ values (where $\alpha_{game} \in [1,10000]$), a value of $\widehat{\theta}_i$ is produced for each of the $\alpha_{game}$ in the grid.
-For each of the $\widehat{\theta}_i$ that are estimated, we'll use the section that wasn't in the training set, the RMSE is calculated here by: \hh{XXX I am not sure this really is the RMSE - where did the squaring and the square rooting go?}
+For each of the $\widehat{\theta}_i$ that are estimated, we'll use the section that wasn't in the training set, the RMSE is calculated here by:
 \begin{equation}
-\text{RMSE} = \frac{ \sum r_{u,i}^{(1)} - (\theta_g + \theta_i)}{n}
+\text{RMSE} = \left(\frac{ \sum (r_{u,i}^{(1)} - (\theta_g + \theta_i))^2}{n}\right)^{\frac{1}{2}}
 \end{equation}
@@
 @
-\caption{\label{fig:k} \hh{You still need a caption here.}}
+\caption{\label{fig:k} RMSE of candidate k values. For this player, 18 is the optimal value.}
 \end{figure}
 
-\section{Technical Implementation} \hh{give references for R (to be a nice R user), shiny, and the other packages that you are using.
-citation() and citation("packagename") makes your life easy with that.}
-This algorithm is implemented in R  using shiny as a vehicle for a simple, easy to use user interface.
+\section{Technical Implementation} 
+This algorithm is implemented in R \citep{rcite}  using shiny \citep{shiny} as a vehicle for a simple, easy to use user interface.
+Additionally, the packages RSQLite \citep{RSQLite}, dplyr \citep{dplyr}, and magrittr \citep{magrittr} were used.
 Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that will be useful in the classifier aggregation and k-nearest-neighbors section; the predicted ratings table, which will store predicted ratings if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user.
@@
 @
-\caption{\label{fig:subdplot} A player preference plot for subdomains for the user cwmassey.  \hh{XXX fix the y label, if showing it}}
+\caption{\label{fig:subdplot} A player preference plot for subdomains for the user cwmassey.}
 \end{figure}

### 2015-04-08 paper.Rnw
any future tense removed, no contractions, some typos

@@
 A few methods of varying usefulness already exist.
-Boardgamegeek.com, a popular site among frequent board gamers, has a recommendation system that allows users to find a particular game they like, and then from there it will recommend more games that a person may like.
+Boardgamegeek.com, a popular site among frequent board gamers, has a recommendation system that allows users to find a particular game they like, and then from there it  recommends more games that a person may like.
 A pretty thorough analysis of it (and its issues) can be found at \url{http://boardgamegeek.com/wiki/page/Game_Recommendation_Algorithm}.
@@
 
-Recommendations for the individual aren't easy to come by. There are a few automated systems out there (such as the boardgame 
+Recommendations for the individual are not   easy to come by. There are a few automated systems out there (such as the board game 
 recommender bot on reddit), but for the most part they are time intensive and often lack interactivity or any sort of information that would be useful to the user beyond a few single games that it might recommend.
 
-In this paper, we seek to build a system the will recommend board games to users based on their own recommendations as well as the recommendations of others.
-This system will offer the user not only recommendations for a very large number of games, but it will also offer insights to the user about what sorts of games they like, not only scores for particular games.
+In this paper, we seek to build a system that recommends board games to users based on their own recommendations as well as the recommendations of others.
+This system offers the user not only recommendations for a very large number of games, but it  also offers insights to the user about what sorts of games they like, not only scores for particular games.
 
@@
 Users can create an account and then rate games they've played as well as keep an inventory of games they own.
-All of this information is accessible and scrape-able, and will be the data that drives our analysis.
+All of this information is accessible and scrape-able, and is the data that drives our analysis.
 
-There are two types of particular information that we will be interested in: user ratings of the games and characteristics about the game (referred to as classifiers in the modeling section).
+There are two types of particular information that we are interested in: user ratings of the games and characteristics about the game (referred to as classifiers in the modeling section).
 The characteristics we're interested in fall into these categories as outlined in table~\ref{tab:categories}.
@@
 we make sure to check that all of their current information is correct.
-We look through known ratings and  scrape  their userpage on BoardGameGeek.
+We look through known ratings and  scrape  their user page on BoardGameGeek.
 Based on this, we  revise old ratings and add new ones, as well as check to see that all of the games that this user has rated are included in the database!
@@
 
-Our approach is to borrow from a single element of their shotgun modeling. Where they use many different models of many different types, we will choose a simple implementation of a single model.
+Our approach is to borrow from a single element of their shotgun modeling.
+Where BigChaos and BellKor use many different models of many different types, we choose a simple implementation of a single model.
 
 There are a few things that make sense to model first.
-Certain games are more popular than other games, and certain people will naturally be more or less generous than others in their ratings.
-These two effects are very important to measure, as the first will allow us to provide estimates for people that have rated no games (and therefore we know nothing about), and the second will allow us to fine tune estimates for people who have rated many games.
-We will implement these in a similar manner to the way that BigChaos implemented the Global Effects model \citep{bigchaos}.
-While they implemented 14 steps including Movie Effect, User Effect, and those variables crossed with variables like time since rating, average rating for the movie, production year, standard deviations, as well as others, we'll stick to only using 2 steps: an effect for each game and an effect for each player.
+Certain games are more popular than other games, and certain people are naturally more or less generous than others in their ratings.
+These two effects are very important to measure, as the first allows us to provide estimates for people that have rated no games (and therefore we know nothing about), and the second  allows us to fine tune estimates for people who have rated many games.
+We  implement these in a similar manner to the way that BigChaos implemented the Global Effects model \citep{bigchaos}.
+While they implemented 14 steps including Movie Effect, User Effect, and those variables crossed with variables like time since rating, average rating for the movie, production year, standard deviations, as well as others, we   stick to  using only two steps: one effect for each game and one effect for each player.
 To compare models to each other we use root mean square error (RMSE).
@@
 From the RMSE of the probe set of the Netflix data, the difference between using only user/movie effects and using all 14 effects was .02, a decrease of about 2\%.
-Noting this negligible increase in accuracy and recognizing my lack of significant computing resources, we will only use user and game effects.
-My model is fit on my home desktop, which has an Intel Core i7-4770K CPU @ 3.5GHz and 32GB of RAM running 64-bit Windows 8.
-With the help of additional computing cores, RAM, and better parallelization, we could hope to achieve the 2\% accuracy increase seen in the Netflix Data, but more than likely it wouldn't be worth it, as a quick cost/benefit analysis reveals that I don't need to hit the nail on the head every time, I just need to implement a method that reliably works a good enough percentage of the time.
-To this coder, a 2\% increase in accuracy just isn't worth the money.
+Noting this negligible increase in accuracy and recognizing my lack of significant computing resources, we  only use user and game effects.
+This model is fit on my home desktop, which has an Intel Core i7-4770K CPU @ 3.5GHz and 32GB of RAM running 64-bit Windows 8.
+With the help of additional computing cores, RAM, and better parallelization, we could hope to achieve the 2\% accuracy increase seen in the Netflix Data, but more than likely it would not   be worth it, as a quick cost/benefit analysis reveals that I do not   need to hit the nail on the head every time, I just need to implement a method that reliably works a good enough percentage of the time.
+To this coder, a 2\% increase in accuracy just is not   worth the money.
 
@@
 This kills two birds with one stone.
-We'll create, for each user, a rating for each game mechanic and genre.
+We  create, for each user, a rating for each game mechanic and genre.
 Then, since there are so many mechanics and genres of game, a player only needs to rate a small, albeit diverse set of games before we can correlate that player's habits with other players, even if they've never played any of the same games!
-Then we can use those to filter rating results, as we'll know what board game mechanics that a particular player will enjoy.
+Then we can use those to filter rating results, as we   know what board game mechanics that a particular player  enjoy.
 
-We now have a very simple two step model that will allow us to predict ratings for new board games that a person may not have even played!
+We now have a very simple two step model that  allows us to predict ratings for new board games that a person may not have even played!
 Also, it allows us to rate games that nobody has played before: We can use the overall mean as a starting point, then for an individual player, we can adjust the rating based on their ratings of similar games from other genres.
@@
 Here, we cover the specifics of both.
-We will step through the model in sequence, removing the effects that games and players naturally place on ratings (some games are naturally rated higher than others, some players are more or less generous than others).
-After removing those larger effects, we'll model the remaining nuance using characteristics about the games that were rated.
-From there, we'll strive to find users who have similar tastes in game characteristics, and then finally use that closeness to find ratings for games that a user has not yet rated.
+We  step through the model in sequence, removing the effects that games and players naturally place on ratings (some games are naturally rated higher than others, some players are more or less generous than others).
+After removing those larger effects, we   model the remaining nuance using characteristics about the games that were rated.
+From there, we   strive to find users who have similar tastes in game characteristics, and then finally use that closeness to find ratings for games that a user has not yet rated.
 
@@
 
-We'll estimate these effects one at a time using a hierarchical structure.
+We   estimate these effects one at a time using a hierarchical structure.
 Define $r_{u,i}^{(1)}$ be the rating from user $u$ for game $i$.
-The first effect we'll fit is a global mean, which we'll denote with the parameter $\theta_g$, and we'll model the ratings using the following formula:
+The first effect we   fit is a global mean, which we   denote with the parameter $\theta_g$, and we   model the ratings using the following formula:
 
@@
 $\theta_g$ is estimated using a simple mean of the $r_{u,i}^{(1)}$'s.
-After removing a global mean, we'll model the global effect for each game using the residuals from the global mean step.
-For this purpose, define $r_{u,i}^{(2)} = \epsilon_{u,i}^{(1)}$, which we'll model using global effect for each game, denoted $\theta_i$.
+After removing a global mean, we   model the global effect for each game using the residuals from the global mean step.
+For this purpose, define $r_{u,i}^{(2)} = \epsilon_{u,i}^{(1)}$, which we   model using global effect for each game, denoted $\theta_i$.
 The residuals are then modeled using the following, where $x_i$ is an indicator variable that the residual belongs to game $i$:
@@
 %  The general idea is to remove the global mean from all ratings to get a residual for each observation.
-We'll group these residuals by game, and then find an effect for that game.
-Subtracting this effect from the residual, we'll take those residuals and repeat the same for each user.
+We will  group these residuals by game, and then find an effect for that game.
+Subtracting this effect from the residual, we will  take those residuals and repeat the same for each user.
 The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference, \hh{you still need to paraphrase the motivation here.}).
@@
 % 
-% However, we'll approach this in steps, as follows: \hh{what is the benefit of using a hierarchical approach?}
+% However, we will  approach this in steps, as follows: \hh{what is the benefit of using a hierarchical approach?}
 % 
@@
 % 
-% And we'll use the estimation method outlined in the paper (need ref) for each, as follows below:
+% And we will  use the estimation method outlined in the paper (need ref) for each, as follows below:
 % 
@@
 For example, the game Dragonmaster was published in 1981, has a playing time of 30 minutes, is under the family ``Animals: Dragons", falls under the categories ``Card Game", ``Fantasy Game", is in the subdomain ``Strategy Games", uses the mechanic ``Trick-taking", and seats 3 or 4 players.
-Each one of these will be treated as a classifier for the game.
+Each one of these is treated as a classifier for the game.
 A table storing this information by GAME\_ID can be found in Table 2.
 
-In order to model the remaining residuals, we'll give each player a score based on how they rated games with similar classifiers; that is, how did they rate games with the ``Trick-taking" mechanic, or how did they rate ``Card Games"?
+In order to model the remaining residuals, we   give each player a score based on how they rated games with similar classifiers; that is, how did they rate games with the ``Trick-taking" mechanic, or how did they rate ``Card Games"?
 
@@
 For example, if the classifier was the mechanic ``Trick-taking" and the user was $u = 3$, the classifier score would be the mean of all of the $\epsilon_{3,i}^{(3)}$ for \textit{all} games that the user has rated that have the ``Trick-taking" mechanic.
-If its the case that a user has rated no games with a particular classifier, we'll leave it as an unobserved value.
+If it is the case that a user has rated no games with a particular classifier, we   leave it as an unobserved value.
 
@@
 Now we have a few classifier values defined for each user who has rated at least one game.
-In order to create predictions for other games that the user hasn't played, we'll find the correlation between players based on their classifier values.
+In order to create predictions for other games that the user has not   played, we   find the correlation between players based on their classifier values.
 For each pair of players, the correlation between them can be found by finding all of the classifiers that have scores for both players, and then finding the correlation between those values.
-To make sure that we do not find high correlations between players who only have very few classifier scores in common, we'll on consider a correlation between players if they have more than 6 common classifier scores.
+To make sure that we do not find high correlations between players who only have very few classifier scores in common, we consider a correlation between players, if they have more than six common classifier scores.
 
-Now to generate an estimate for user $u$ for game $i$, we'll find all of the players who have a correlation value with user $u$ who have also rated game $i$.
+Now to generate an estimate for user $u$ for game $i$, we   find all of the players who have a correlation value with user $u$ who have also rated game $i$.
 The rating for game $i$ is then the weighted average of the $k$ most highly correlated (or nearest) neighbors of user $u$.
 
-The weights for this will be $(1 - \rho_{u,k})^{-1}$ where $\rho_{u,k}$ is the correlation between user $u$ and user $k$.
-The number of neighbors to use $k$ will be determined by cross-validation on a user-to-user basis.
-In the end $z_{u,i}$ will be the weighted average of the k-nearest-neighbors scores for user $u$ and game $i$, as given in the equation below (where $u_{[j]}$ is the user with the $j^{\text{th}}$ highest correlation between themselves and user $u$):
+The weights for this are $(1 - \rho_{u,k})^{-1}$ where $\rho_{u,k}$ is the correlation between user $u$ and user $k$.
+The number of neighbors to use $k$ is determined by cross-validation on a user-to-user basis.
+In the end $z_{u,i}$ is the weighted average of the k-nearest-neighbors scores for user $u$ and game $i$, as given in the equation below (where $u_{[j]}$ is the user with the $j^{\text{th}}$ highest correlation between themselves and user $u$):
 
@@
 There are three parameters that the model relies on: The two $\alpha$ values that balance out biasedness in the global effects section, and the $k$ in the k-nearest neighbors portion of the algorithm.
-The $\alpha$'s will be done globally (that is, every individual will use the same values for $\alpha_{game}$ and $\alpha_{player}$), while the values for $k$ will be done locally (each user will have their own value of $k$).
+The $\alpha$s  are fitted globally (that is, every individual uses the same values for $\alpha_{game}$ and $\alpha_{player}$), while the values for $k$ are done locally (each user has their own value of $k$).
 
@@
 Since the game parameter is fit first, we fit the $\alpha$ tuning parameter for the game effect first.
-After finding an optimal value, we'll then fit the $\alpha$ tuning parameter for the player effect, as the game effect is needed to estimate the player effect.
+After finding an optimal value, we fit the $\alpha$ tuning parameter for the player effect, as the game effect is needed to estimate the player effect.
 
-The optimization will use ten-fold cross validation, wherein the data is randomly divided into 10 separate sections of roughly equal size.
-Then for each of the 10 sections, we'll treat the section as a test set and the other 9 to use as a training set.
+The optimization uses ten-fold cross validation, wherein the data is randomly divided into 10 separate sections of roughly equal size.
+We then treat each of the 10 sections as a test set and the other 9 to use as a training set.
 $\widehat{\theta}_g$ is fit first, and then from a grid of $\alpha_{game}$ values (where $\alpha_{game} \in [1,10000]$), a value of $\widehat{\theta}_i$ is produced for each of the $\alpha_{game}$ in the grid.
-For each of the $\widehat{\theta}_i$ that are estimated, we'll use the section that wasn't in the training set, the RMSE is calculated here by:
+For each of the $\widehat{\theta}_i$ that are estimated, we   use the section that was not   in the training set, the RMSE is calculated here by:
 \begin{equation}
-\text{RMSE} = \left(\frac{ \sum (r_{u,i}^{(1)} - (\theta_g + \theta_i))^2}{n}\right)^{\frac{1}{2}}
+\text{RMSE} = \sqrt{\frac{1}{n} \sum \left(r_{u,i}^{(1)} - (\theta_g + \theta_i)\right)^2}
 \end{equation}
@@
 where we sum over all rated games in the test set, and $n$ is the number of ratings in the test set.
-The value of $\alpha_{game}$ that has the lowest RMSE will be used.
-in figure~\ref{fig:gamealpha}, we see the optimal value for $\alpha_{game}$ sits in a nice trough, with the minimum at $\alpha_{game} = 1.802$.
+The value of $\alpha_{game}$ that minimizes the overall RMSE is used.
+In figure~\ref{fig:gamealpha}, we see the optimal value for $\alpha_{game}$ sits in a nice trough, with the minimum at $\alpha_{game} = 1.802$.
 
@@
 After finding the optimal value for $\alpha_{game}$ we can move on to optimizing $\alpha_{player}$.
-Once again, the data will be divided randomly into approximately 10 equally sized sections, and each will be treated as an individual test set while the other 9 combined will serve as a training set.
-A grid of $\alpha_{player} \in  [1,10000]$ will be used to create a value for $\widehat{\theta}_{u}$, and for each of those values, we'll calculate the RMSE as below:
+Once again, the data is divided randomly into approximately 10 equally sized sections, and each is treated as an individual test set while the other 9 combined will serve as a training set.
+A grid of $\alpha_{player} \in  [1,10000]$ is used to create a value for $\widehat{\theta}_{u}$, and for each of those values, we   calculate the RMSE as below:
 
@@
 
-Once again, the value of $\alpha_{player}$ with the lowest RMSE will be the value used, which as can be seen in ~\ref{fig:playeralpha} is $\alpha_{player} = 2.022$.
+Once again, the value of $\alpha_{player}$ with the lowest RMSE is the value used, which as can be seen in ~\ref{fig:playeralpha} is $\alpha_{player} = 2.022$.
 \begin{figure}
@@
 Additionally, the packages RSQLite \citep{RSQLite}, dplyr \citep{dplyr}, and magrittr \citep{magrittr} were used.
-Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that will be useful in the classifier aggregation and k-nearest-neighbors section; the predicted ratings table, which will store predicted ratings if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user.
-This data storage will allow for ease of implementation: we only really need the ratings table and the classifier table to do all the work, and the key tables are really only needed to display the information to the user in a format which they will understand.
+Behind the scenes, the data is stored in a data base with six tables: A player key table, which relates the players information to a unique player userid; A game key table, which relates game information to a unique game id; A ratings table which stores the ratings (a score from 1-10) along with the userid of the person who rated it and the game id which the rating is for; the classifiers table, relates the game id to characteristics about the game that is useful in the classifier aggregation and k-nearest-neighbors section; the predicted ratings table, which  stores predicted ratings if ratings have already been fit for a particular user; and the aggregated classifier ratings for each user.
+This data storage  allows for ease of implementation: we only really need the ratings table and the classifier table to do all the work, and the key tables are really only needed to display the information to the user in a format which they  understand.
 
-In an implementation of this, the application will begin by having the user enter the username they've registered on \url{boardgamegeek.com}.
+In an implementation of this, the application  begins by having the user enter the username they've registered on \url{boardgamegeek.com}.
 As seen in (~\ref{fig:blanksc}), there is an area to enter the users information in, and a button just below that, when clicked, begins the process.
@@
 \end{center}
-Instead of logging the users ratings directly through the application, we will instead leverage the existing infrastructure of boardgame geek.
-After checking that its a valid username, the application will scour boardgamegeek to see if the database's information is complete: that is, do we have this user's data in the player key?
-Have they rated games that aren't in our database?
+Instead of logging the users ratings directly through the application, we  leverage the existing infrastructure of board game geek.
+After checking that its a valid username, the application  scours {\tt boardgamegeek} to see if the database's information is complete: that is, do we have this user's data in the player key?
+Have they rated games that are not   in our database?
 Have they re rated games that are in our database?
-Have they rated games that aren't in our database?
-If so, we also need to bring in other users ratings for any games that we may not have had in our database (since a single rating for a game isn't very useful and will greatly skew the results for that game).
-After this initial step and database update, we can see if they already have saved predicted ratings in the system, and if not, we'll proceed to model fitting.
+Have they rated games that are not   in our database?
+If so, we also need to bring in other users ratings for any games that we may not have had in our database (as a single rating for a game is not   very useful and  greatly skews the results for that game).
+After this initial step and database update, we can see if they already have saved predicted ratings in the system, and if not, we   proceed to fit the model.
 
 We first apply the global effects strategy as described in the modeling section.
-The optimization of the $\alpha$ tuning parameters will have occurred sometime before this and the algorithm will use use those stored values.
+The optimization of the $\alpha$ tuning parameters will have occurred sometime before this and the algorithm  uses those stored values.
 A weekly or monthly automatic updating procedure is to be used to keep these values up to date as more data comes into the system.
@@
 Finally, after the 10-fold cross-validation, we're ready to run the other games through the algorithm to get a predicted score.
-The games that the user has already rated will then be replaced with those ratings, and the full list will be outputted for the user in the "My Recommendations" Tab, which can be seen in Figure ~\ref{fig:act_use}.
+The games that the user has already rated will then be replaced with those ratings, and the full result is listed for the user in the ``My Recommendations" Tab, which can be seen in Figure ~\ref{fig:act_use}.
 \begin{figure}\centering
 \includegraphics[scale=0.4]{active_use.png}
-\caption{\label{fig:act_use} The "My Recommendations" tab, populated for user cwmassey.}
+\caption{\label{fig:act_use} The ``My Recommendations" tab, populated for user cwmassey.}
 \end{figure}
@@
 They can then search by their preferred subdomain, and get ratings for games that fit that description.
-This plot is provided to the user within the "My Category Ratings" tab.
+This plot is provided to the user within the ``My Category Ratings" tab.
 The plot for the specific category rating can be chosen by the user, so the user can select from Category, Family, Mechanic, Subdomain, Number of Players, Playing Time, or Year Published.
@@
 
-If a user wants to rate a new game, they have to go back to BoardGameGeek and rate games there. The app will then need to be restarted, which will pull the new rating, and update the database.
+If a user wants to rate a new game, they have to go back to BoardGameGeek and rate games there. The app  then needs to be restarted, which will pull the new rating, and update the database.
 
 \section{Conclusion}
-While the result was not as smooth as hoped, the Netflix recommender system made an excellent base off of which to build a recommender system for board games.
-It would not be too much of a stretch to imagine that, if you had a reasonable amount of user rating data about something, and you had enough descriptive data about that thing, that a similar implementation would be of great use there as well.
+While the result was not as smooth as hoped, the Netflix recommender system made an excellent base for building a recommender system for board games.
+It would not be too much of a stretch to imagine that, with a reasonable amount of user rating data about something, and enough descriptive data about that thing, that a similar implementation would be of great use there as well.
 
@@
 The first of which is to speed up the prediction generating process.
-Currently, each time a user starts the app with new data, the new data is fetched and the whole model is fit again, including generating a truly massive correlation matrix.
+Currently, each time a user starts the app with new data, the whole model is fit again, including generating a truly massive correlation matrix.
 Instead of doing this each time a person logs in, it may be more advantageous to fit the correlation matrix daily, and then only update the global effects portion of the model for each user, as that portion is relatively snappy.
 Hopefully this would bring the fitting time down from 20 minutes to a matter of seconds.
-Additionally, we don't have the full data that is contained on BoardGameGeek.
+Additionally, we do not   have the full data that is contained on BoardGameGeek.
 There is much more data to scrape than time for this paper allowed, so going and continually updating the database would be advantageous.

### 2015-04-08 paper.Rnw
've contraction

@@
 For data, we turn to \url{boardgamegeek.com}, a website that holds information, ratings, and reviews for over 75,000 games.
-Users can create an account and then rate games they've played as well as keep an inventory of games they own.
+Users can create an account and then rate games they have  played as well as keep an inventory of games they own.
 All of this information is accessible and scrape-able, and is the data that drives our analysis.
@@
 In addition, we'd also like to compare a users ratings to other users, and hopefully use the behavior of similar users to predict new ratings for someone.
-A tempting first step is to, for a particular game, find other people who have rated that game who have rated games that you've rated.
+A tempting first step is to, for a particular game, find other people who have rated that game who have rated games that you have  rated.
 Then, you could find people who rated games similarly to you, and use their ratings of the new game as an estimate.
@@
 We  create, for each user, a rating for each game mechanic and genre.
-Then, since there are so many mechanics and genres of game, a player only needs to rate a small, albeit diverse set of games before we can correlate that player's habits with other players, even if they've never played any of the same games!
+Then, since there are so many mechanics and genres of game, a player only needs to rate a small, albeit diverse set of games before we can correlate that player's habits with other players, even if they have  never played any of the same games!
 Then we can use those to filter rating results, as we   know what board game mechanics that a particular player  enjoy.
@@
 \subsubsection{Back to Rating}
-From what we've produced above in equations (2)-(9), of the estimated rating for user $u$ and game $i$ is given by:
+From what we have  produced above in equations (2)-(9), of the estimated rating for user $u$ and game $i$ is given by:
 
@@
 
-In an implementation of this, the application  begins by having the user enter the username they've registered on \url{boardgamegeek.com}.
+In an implementation of this, the application  begins by having the user enter the username they have  registered on \url{boardgamegeek.com}.
 As seen in (~\ref{fig:blanksc}), there is an area to enter the users information in, and a button just below that, when clicked, begins the process.
